### 2019-01-02 content/post/2019-01-02-transit.Rmd

---
title: Transit
author: Effi Feldblum
date: '2019-01-02'
slug: transit
categories: []
tags: []
---

Please please work

### 2019-01-02 public/post/2019-01-02-transit.Rmd

---
title: Transit
author: Effi Feldblum
date: '2019-01-02'
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      results='show', cache=FALSE, autodep=FALSE)

knitr::opts_knit$set(root.dir = 'data/transit/')
```

```{r, include = FALSE, message= FALSE, warning = FALSE}
# include all packages here if nec.
library(kableExtra)
library(tigris)
library(tidyverse)
library(leaflet)
library(sp)
library(gridExtra)
library(DT)
library(data.table)
library(rowr)

```

```{r shapefiles, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}
philly <- tracts(state = "PA", county = "Philadelphia")
dc <- tracts(state = "District of Columbia")
ny <- tracts(state = "NY", 
             county = c("New York County", "Bronx County",
                        "Kings County", "Queens County", "Richmond County")  )

```

```{r dataload, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}

load("GET_url_test.RData")
load("content_url_test.RData")
load("final_url_test.RData")
load("json_url_test.RData")

load("googleway_raw_data.RData")

load("philly_300_points.RData")

load("aplay_PH.RData")
load("aplay_DC.RData")
load("aplay_NY.RData")

```

```{r base_ggplots, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}
gg_ph <- 
  ggplot() +
  geom_polygon(data = philly, aes(long, lat, group = group), 
               fill = "#051633", color="#051633") +
  coord_map() +
  labs(x = NULL, y = NULL) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = '#e8edf7'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

gg_ny <- 
  ggplot() +
  geom_polygon(data = ny, aes(long, lat, group = group), 
               fill = "#051633", color="#051633") +
  coord_map() +
  labs(x = NULL, y = NULL) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = '#e8edf7'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

gg_dc <- 
  ggplot() +
  geom_polygon(data = dc, aes(long, lat, group = group), 
               fill = "#051633", color="#051633") +
  coord_map() +
  labs(x = NULL, y = NULL) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = '#e8edf7'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

```

# Part 1: The Idea  {-}

I was recently in a car with one of my best friends and his new wife[^1] (also a good friend of mine). We were making our way back from our parents joint 4th of July BBQ in Philadelphia towards Washington DC where we had all attended University. These two friends of mine were coming back down to prepare for their fast approaching move to NYC [^2].

Car talk moved towards talking about different public transit systems we've used both around the country and world. Being poor students and recent graduates and all moving in the near future, we are three of us particulary interested in our city's system. This was, of course, mostly complaining about the systems and comparing their negatives. This led me to an, I think, interesting question:

If you were to be dropped in a city at a random point within that cities limit and given another random point to get to, again within the city limits, which city would this be the easiest in?

I will here be trying to begin answering that question. In order to simplify the scope of this exploration, I am limiting myself to looking at New York City, Washington DC, and Philadelphia[^3].

##Plan of Attack 1 {-}
###Packages  {-}
This analysis will be occuring entirely in R, so the first step is to load up some necessary packages to get started.

###Let's Get Started {-}

The first idea I had in solving this was to create a box around each city, get the longitude-latitude codes of the two corners, then get random points using `points <-  runif(num_points, lower_long/lat_barrier, upper_long/lat_barrier)`.

As an example:
```{r}
x <- data.frame( x= runif(3, 1,50), y = runif(3, 51, 100))
```
```{r, echo=FALSE}
x %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
                position = "float_right")
```

Runif will return to us a specified set of random numbers (here `3`) from a uniform distribution. It is a similar function to RandBetween in Excel. You can see the results in the table to the right.

So, I needed to get approximated bounded corners for the city. The way I decided to do this was to use googlemaps functionality of finding longitude and latitude of any point clicked.

```{r echo=FALSE, out.width = '100%', eval = FALSE}
knitr::include_graphics("google_maps_whats_here.png")

```
```{r echo=FALSE, out.width = '35%', eval = FALSE}
knitr::include_graphics("Untitled.png")
```

<br>

So, after some slight maneuvering, we can use the two sets of coordinates to create this:
```{r echo = FALSE}
leaflet(philly) %>%
    addProviderTiles(providers$Stamen.Toner) %>%
  addRectangles(
    lng1=-75.381713, lat1=40.143756,
    lng2=-74.949813, lat2=39.885019, fillColor  = "transparent"
  )%>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")

```

Great! Okay, so we can use our coordinate codes of `-75.381713, 40.143756` and `-74.949813, 39.885019` within our `runif` function. As a test, I'll generate 100 random points and map them out. Below I've provided a sample of the points in a table.

```{r}
longitude <-runif(100, -75.381713, -74.949813)
latitude <- runif(100, 39.885019, 40.143756)
my_df <- data.frame(longitude, latitude)
```

```{r, echo=FALSE}
my_df[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE,
                position = "float_right")
```

```{r echo = FALSE}

leaflet(philly) %>%
  addProviderTiles(providers$Stamen.Toner) %>%
  addRectangles(
    lng1=-75.381713, lat1=40.143756,
    lng2=-74.949813, lat2=39.885019, fillColor  = "transparent"
  )%>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addCircleMarkers(data=my_df,radius =1.5,
                   color= "#ff0505")

```

## A Clear Issue {-}
This is obviously an issue.  
Cities are not specificially oriented in the plot of the Earth's coordinate system, they are a "random" human choice. Even worse (becuase with motivation, the perspective of the graph can be changed), cities are weird shapes. Humans shaped these amazing metropolises that span whatever land they decide. A very cool thought, but highly inconvenient for me.

# Part 2: Shapefiles and Progress {-}

```{r echo=FALSE, out.width = '50%',eval = FALSE}
knitr::include_graphics("Septa_logo.png")

```

The main lesson from part 1 is that I'm going to have to be a little less lazy in this project. The minitua is that I needed:

   1. a way to graph only the city limits itself  
   2. a way to generate random points within that irregular polygon

Googling told me that what I needed was a shapefile, "The shapefile format is a popular geospatial vector data format for geographic information system (GIS) software...The shapefile format can spatially describe vector features: points, lines, and polygons, representing, for example, water wells, rivers, and lakes."[^4]

At first, it looked like this required me to download a bunch of files from the internet and load them all into R. At first because of the blessed package Tigris which allows users to download shapefiles from the US Census Bureau.

The below code is using the TIGRIS package to grab the files for Philadelphia, New York City, and Washington D.C.
```{r echo = FALSE, message= FALSE, warning = FALSE, fig.width=10, fig.height=5}

grid.arrange(gg_ph, gg_ny, gg_dc, ncol = 3)

```
Okay great - we have our city polygons.
## Getting Random Points {-}
To work with the shapefiles in R, the package `sp` was used. This package has a great function `spsample` which magically spits out the random numbers. At first, I still thought I'd have to figure out how to divide the cities into shapes I could find random points within then recombine.

I thought a good starting point would be to get 300 points, 150 origin and 150 destination, and find the directions between them. I'll show an example of that workflow using only Philadelphia.

```{r, cache = TRUE}
number_sample <- 150  #This is the number of trips we want to test
city_file <- philly #This is what city to get points for

origin <- spsample(city_file, n = number_sample, "random") %>%
  as.data.frame()
origin <- origin %>%
  select(startLat = y , startLon = x) %>%
  mutate(trip_id = 1:nrow(origin))

destination <- spsample(city_file, n = number_sample, "random") %>%
  as.data.frame()
destination <- destination %>%
  select(endLat = y , endLon = x) %>%
  mutate(trip_id = 1:nrow(destination))

#To combine the origins and destinations
trip <- left_join(origin,destination,by = "trip_id") %>%
  select(trip_id,startLat, startLon, endLat, endLon)
```

```{r, echo=FALSE,warning = FALSE}
trip1 <- trip %>%
  select("start latitude" = startLat, "start longitude" = startLon,  "end latitude" = endLat, "end longitude" = endLon )

datatable(trip1, options = list(
  dom = 't'
))
```
<br>

```{r, echo = FALSE}
origin <- origin %>%
  select(latitude = startLat, longitude = startLon, trip_id)%>%
  mutate(type = rep("origin"))
destination <- destination %>%
  select(latitude = endLat, longitude = endLon, trip_id)%>%
  mutate(type = rep("destination"))

loc_both <- rbind(origin, destination)

city_origin_destination <- gg_ph + 
  geom_point(data = loc_both, aes(longitude, latitude, group = type, color =type), size =  1.5)+
  scale_color_manual(values=c("#db7918", "#1dbdf7"))

city_origin_destination  
```

Now if we want to view the "routes" themselves if the one went as the bird flew.

```{r, echo = FALSE}

trip_a2b_lines <- city_origin_destination +
  geom_path(data=loc_both, aes(x=longitude, y=latitude, group=trip_id), size=0.5, alpha=0.5, color = "white")

trip_a2b_lines
```

##Getting Routes {-}

So now we have this shit.
great package
google way // attaches to google api

```{r, echo = FALSE}
#loading a predone dataframe so that the document doesn't
#ping the API everytime I reknit...or ever really

```

```{r, eval = FALSE}
googleway_raw_data <- lapply(1:nrow(trip), function(x){

  google_distance(origin = (locations[x, 2:3]),
                  destination = (locations[x, 4:5]),
                  key = key,
                  mode = "transit")

})

```
This results in a list of lists where each holds the data of 1 trip.
```{r}
googleway_raw_data[[5]] #looking at the fifth list, which is also the fifth set of coordinates
```

So, some serious data munging is needed... *cue the montage music*

```{r, warning= FALSE}
foo <- do.call(what = "rbind",
               args = lapply(googleway_raw_data, as.data.frame))

foo <-  separate(foo, elements, into = c("km", "value", "duration", 
                    "value2", "currency", "cost", "uk", "status2"), ",")

foo <- 
  select(foo, destination_addresses, origin_addresses, km, duration, cost, status)
head(foo)
```

Now to get rid of all that annoying fluff[^5]

```{r}
##source of numextract function: http://stla.github.io/stlapblog/posts/Numextract.html
numextract <- function(string){ 
  str_extract(string, "\\-*\\d+\\.*\\d*")
}

foo$km <- lapply(foo$km, numextract)%>%
  as.numeric()
foo$cost <- substring(foo$cost, 11)
foo$cost <- substr(foo$cost,1,nchar(foo$cost)-1) %>%
  as.numeric()
foo$duration <- substring(foo$duration, 26)
foo$duration <- substr(foo$duration,1,nchar(foo$duration)-1)

#Splitting up duration from XX hour(s) XX minute(s), to seperate columns

foo1 <- separate(foo, duration, into = c("hours", "minutes"), sep = -7)
foo1$hours <- gsub("[^0-9]", "", foo1$hours)
foo1$minutes <- gsub("[^0-9]", "",foo1$minutes)
foo1$hours <- as.numeric(foo1$hours)
foo1$minutes <- as.numeric(foo1$minutes)

#for trips that take less than an hour
#doesnt murk with invalid trips because of other apparent NAS
foo1$hours[is.na(foo1$hours)] <- 0

#creating a total minutes section
foo1 <- foo1 %>% 
  mutate(duration_minutes = (hours*60)+minutes)%>%
  mutate(trip_id = 1:nrow(foo1))

#merging with trip to have coordinates
full_AB_GOOGLE <- left_join(foo1, trip,by = "trip_id")

```

```{r, echo=FALSE,warning = FALSE}
full_AB_GOOGLE1 <- full_AB_GOOGLE

full_AB_GOOGLE1$startLat <- round(full_AB_GOOGLE1$startLat, digits = 4)
full_AB_GOOGLE1$startLon <- round(full_AB_GOOGLE1$startLon, digits = 4)
full_AB_GOOGLE1$endLat <- round(full_AB_GOOGLE1$endLat, digits = 4)
full_AB_GOOGLE1$endLon <- round(full_AB_GOOGLE1$endLon, digits = 4)
full_AB_GOOGLE1 <- full_AB_GOOGLE1 %>%
  select(origin_addresses, destination_addresses, duration_minutes, km, cost, startLat, startLon, endLat, endLon)

datatable(full_AB_GOOGLE1, options = list(
  dom = 'ltp'
))
```

## Our first "good" data {-}
Let's take a quick peak at what the data tells us before moving on.

```{r warning = FALSE}
#selecting a few choice variables and adding miles
foo <- full_AB_GOOGLE %>%
  select(trip_id, kilometers = km, duration = duration_minutes, cost)%>%
  mutate(miles = round(kilometers/1.609344, digits =1))

miles <- as.data.frame(c(summary(foo$miles)))
duration <-as.data.frame(c(summary(foo$duration)))
summary1 <- cbind(miles, duration)%>%
  round(digits =2)
colnames(summary1) <- c("miles", "durations")

summary1 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
                position = "float_right")

```

```{r echo = FALSE, message= FALSE, warning = FALSE}

histx <- 
ggplot(data = foo, aes(duration)) + 
  geom_histogram(bins = 10, fill = "#1f4fa3")+
  ggtitle("Histogram of Trip Duration")

scatter <- 
ggplot() +
  geom_point(data =foo, aes(x = duration, y = miles), color = "#f14728")+
  ggtitle("Scatterplot of Duration vs. Miles")

grid.arrange(histx, scatter,  ncol = 2)

```

# Part 3: More Trips {-}

```{r echo=FALSE, out.width = '25%',eval = FALSE}
knitr::include_graphics("mta_logo.png")

```

At this point we've found directions from 150 points to a seperate set of 150 points that were derived from within each city's shapefile. 
This, honestly, seemed a bit weak to me. It answers the question...kind of. The hypothetical situation seemed to call for, well, *more points*.

So the idea is this:

   1. Forget about origins and destinations and combine the 300 points into one list  
   2. Find directions from each point to the other 299 points

```{r warning = FALSE}
#Changing the column names to allow binding
col_names <- c("latitude", "longitude" )

#Getting the columns wanted from origin and destination, then changing the column names
coordinates <- list(origin, destination)%>%
  lapply( function(x) {
  subset(x, select= -trip_id)
  }) %>%
  lapply(setNames, nm = col_names)

#combining the origin and destination coordinate points
loc_both <- rbind(coordinates[[1]],coordinates[[2]])

#splitting each individual set of coordinates into their own dataframe
a <- loc_both %>%
  apply(1, as.data.frame)

#Creating a function that takes each dataframe of a list(here a), replicates the rows the number of times needed, then creates another 2 columns where the other 300 points go (including a duplicate)
multi.lapply <- function(a){
  len <- nrow(loc_both)
  a <- lapply(a, function(x) apply(x, 1, t))
  c <- lapply(a, as.data.frame)
  d <- lapply(c, t)
  e <- lapply(d, function(x)  apply(x, 2, rep, len))
  f <- lapply(e, function(x) cbind(x, 2, loc_both))
  new_col_name <-  c("startLat", "startLon", "trip", "endLat", "endLon")
  g <- lapply(f, setNames, nm = new_col_name)
  h <- bind_rows(g)
  return(h)
}

#Running the created function
philly_300_points <- multi.lapply(a)
philly_300_pointsraw <-philly_300_points[c(3,1,2,4,5)]

#removing the duplicate rows
philly_300_points <- philly_300_pointsraw %>%
  filter(startLat != endLat, startLon != endLon)

head(philly_300_points)
```

So now we have a dataframe of `89700` rows. Which is a subset of the `300*300 = 90,000` observations we expected. This is because each trip has one duplicate row (where the starting longitude = the ending longitude and the starting latitude = the ending latitude).

##Getting the Routes using Google API {-}
As before, we'll be using the googleway package to access the google maps API [^6]. However, I had no interest in letting this function run for the next however much time.

A friend of mine recommended looking into parallel computing over the cloud. It was the perfect recommendation. I setup an AWS account and got an Rstudio environment running, which was a bit of a headache. On July 18, I had finally gotten it up and running. That night I went to sleep and left my computer on as my data was processed.

```{r eval = FALSE}

pls <-  mclapply(1:nrow(tfoo), function(x){

  google_distance(origin = (tfoo[x, 1:2]),
                  destination = (tfoo[x, 4:5]),
                  key = ky,
                  mode = "transit")

})
```

#Part 4: A Catastrophic Issue {-}

```{r echo=FALSE, out.width = '25%', eval = FALSE}
knitr::include_graphics("wmata_logo.png")

```

Okay, I woke up the morning of the July 19th and found the two biggest roadblocks of the project waiting at my computer. What made it castrophic was that, along with these two issues, I had my first set of 90,000 routes.

The first issue was that that day, Google had begun charging for their previously free API. A disastear for my student pockets, and so close to the solution. Well, I was too close to not try and find a replacement.

The second issue was that my use of AWS had cost around $50, not something I was expected. Ooops.

Well let's try and solve the first issue first.

##Finding a Replacement for Google Routes API {-}
Finding another API was only a mild headache. The most annoying aspect was finding one that suppored transit routes and not exclusively driving directions. To the rescue: Here API. It had everything I wanted, except that no kind soul had made an R package to easily access it. It became pretty obvious during that day of research that Here was my best option. I would have to write some of my own functions to get it to work.

###Functions {-}
Basically, I want to take my 3 lists of 90,000 rows and 360,000 indivudal cooridnates and feed them into the Here API.

The API requires that the link I use look like:

> "route.cit.api.here .com/routing/7.2/calculateroute.json?waypoint0=52.5208%2C13.4093&waypoint1=52.5034%2C13.3295&mode=fastest%3BpublicTransport &combineChange=true&app_id=DemoAppId01082013GAL&app_code=AJKnXv84fjrb0KIHawS0Tg"

All coordinates needed to be adjusted to be strings that resemble `52.5208%2C13.4093`.

```{r}
here_mcoord_fix <- function(df){
  latitude <- df[,1]
  longitude <- df[,2]
  url2 <- paste0(longitude, "%2C", latitude)
  return(url2)
}
```

Now, there are several constants in the URL that can be set to variables. Then everything can be easily pasted together into proper URLS.

```{r}
base_url <- "https://route.cit.api.here.com/routing/7.2/calculateroute.json"
xmode <- "&mode=fastest%3BpublicTransport&combineChange=true"
#these two are different per API User
id <- "&app_id=SOME_NUMBERS_AND_LETTERS&"
code <- "app_code=SOME_NUMBERS_AND_LETTERS&departure=2018-07-18T11:00:00-05:00"
```

And here's the function that pulls everything together. It contains the `here_mcoord_fix` function within to cut steps down.

```{r}
get_here_urls <- function(origin, destination){
  xorigin <- here_mcoord_fix(origin)
  xdestination <- here_mcoord_fix(destination)
  z <- paste0(base_url,"?waypoint0=",xorigin, "&waypoint1=",
              xdestination, xmode, id, code, sep="")
  return(z) 
}

```

So let me show the output:
```{r}
origin <- philly_300_points%>%
  select(startLon, startLat)
destination <- philly_300_points%>%
  select(endLon, endLat)

urls <- get_here_urls(origin, destination)

#Let's take a peak at a random 1 of 89700 URLs we've created.
urls[runif(1, 1,89700)]

```

##Running this Code without Emptying my Pockets {-}
So, I decided not to use AWS anymore...maybe if I had deep pockets I would, but I had decided to find an alternative. The first thing I learned was that I did not[^1] really fully understand parallel computing. There were better functions than `mclapply` available. The second thing I realized that I was obsessed with doing the work on the cloud. On one hand, the cloud is definietly better. On the other hand, only if you've got that cash-money.

The solution was to use the `DoParallel` package with computers that have multiple cores, thus performing parallel computations on each core.

###Getting DoParallel Up and Running {-}
```{r, eval = FALSE}
#Setting up parallel with one less core than available on the computer. This is to avoid intense crashing.
no_cores <- detectCores() - 1  
cl <- makeCluster(no_cores)  
registerDoParallel(cl)

#To see if that worked: if it returns 1 - then it didn't work
getDoParWorkers()

#To end the parallel cores
registerDoSEQ()
getDoParWorkers()

```

###Getting the functions ready to access the here API {-}
The first step is getting it running before going parallel and seeing what the outputs look like.

There are some messy looking results so I've tabbed this section to maintain readability. Feel free to look through it if you're interested. Each tab is a view of each output.

#### Everything {-}
The first step is getting it running before going parallel and seeing what the outputs look like:

```{r eval = FALSE}

#We'll use one random url to test
url_test <- urls[runif(1, 1,89700)]
require(httr)
require(jsonlite)

GET_url_test <- GET(url_test)
content_url_test <-  content(GET_url_test, "text")
json_url_test <-   fromJSON(content_url_test, flatten = TRUE)
final_url_test <- as.data.frame(json_url_test)

```

#### Step 1 {-}
```{r}
GET_url_test
```
There are a few good pieces of news in this code:

  *The status is 200 meaning it worked[^7]
  *The data is in JSON format
  *The size is not absurd

#### Step 2 {-}
```{r eval = FALSE}
content_url_test <-  content(GET_url_test, "text")
```

```{r echo = FALSE}
content_url_test
```

#### Step 3 {-}

```{r eval = FALSE}
json_url_test <-   fromJSON(content_url_test, flatten = TRUE)

```

```{r echo = FALSE}
json_url_test
```

#### step 4 {-}
```{r eval = FALSE}
 final_url_test <- as.data.frame(json_url_test)
```

```{r echo = FALSE}
final_url_test

```

###Preparing Looping Functions for Parallel {-}

```{r eval = FALSE}
#gets the Data from HERE API

get_city = foreach(i=urls, .packages='httr') %dopar% {
    GET(i)
  }

status_code(results[[1]]) #Can check some to ensure it worked

#Gets the Json Content
content_city = foreach(i=get_city, .packages='httr') %dopar% {
  content(i, "text")
}

#munging
#makes the json data semi pretty
json_city = foreach(i=content_city, .packages='jsonlite') %dopar% {
  fromJSON(i, flatten = TRUE)
}

df_city = foreach(i=json_city) %dopar% {
    as.data.frame(i)
  }

full <- rbindlist(df_city, fill = TRUE)

```

###Plan of Attack {-}
To solve my 2nd large issue, I decided to leverage the resources of my university. Generally, the idea is to split up the urls into multiple lists, then run each set of URLs independently on several computers.

To split up the URLs:
```{r}
list_of_URLs <- split(urls, ceiling(seq_along(urls)/15000))
str(list_of_URLs)

```

Then on each computer I would run the code, but first specify:
```{r}
urls<- list_of_URLs[[2]] #insert whatever subset you want to use
str(urls)

```

This worked magically. I used several computers over 3 days and got all the data I need. The internet and books constantly preached that 80% of these tasks are data collection / wrangling and only 20% was the actual analysis. I think I finally understand what they mean...But now for the fun! The conclusions and results!!!

# Part 5: Results and Conclusion {-}

## Summary of Results {-}

```{r echo = FALSE, warning=FALSE}
fooph <- play_PH %>%
  group_by(startLat)%>%
  filter(n()<2)%>%
  summarise_all(mean) %>%
  arrange(minutes)

fooDC <- play_DC %>%
  group_by(startLat)%>%
  filter(n()<3 )%>%
  summarise_all(mean) %>%
  arrange(minutes)

foony <- play_NY %>%
  group_by(startLat)%>%
  filter(n()<20 )%>%
  summarise_all(mean) %>%
  arrange(minutes)

acity_row <- data.frame(
DC = c(nrow(play_DC), 90000-nrow(play_DC), nrow(fooDC), 100*((nrow(play_DC)+nrow(fooDC))/90000)),
Philly = c(nrow(play_PH),90000 -nrow(play_PH),nrow(fooph), 100*((nrow(play_DC) + nrow(fooph))/90000)),
NYC = c(nrow(play_NY),90000 -nrow(play_NY),nrow(foony),  100*((nrow(play_DC) + nrow(foony))/90000))
)

rownames(acity_row) <- c("Rows Remaining", "Rows NULL","Got through","Total Percent Removed")

x <- as.matrix(acity_row)

```

```{r echo=FALSE, warning=FALSE}
kable(x, "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right")
```

Before diving into this crazy results chart, I need to admit something...I don't know what I'm talking about when it comes to [GIS](https://en.wikipedia.org/wiki/Geographic_information_system). If I were a real maps whiz, I would have limited my city polygons to only include areas people lived. There were issues with several areas within each city.

  1. Many of these issue areas resulted in, mostly, `NULL` responses from the API I used to get direction routes [^8]. These were easily removed. 
  2. The majority of the rest were a little more *slippery* and got results from the API. An example of one of these were a point in the river would pull `298 NULLS` but would get directions to another random point within the river. These were also easily removed (except for a small few that will be discussed below).
  3. The last bit are a bit more of a question. A question of how evil these made-up gameshow creators are. For example, several points fell into points deep within large philadelphian parks and began with 20-45 minutes of walking. These were kept in[^9].

```{r summary_stats, echo = FALSE, warning = FALSE}
ny_mins_summ <- summary(play_NY$minutes)
ny_mils_summ <- summary(play_NY$miles)

ny_summary_stats <- t(data.frame(ny_mils_summ = as.vector(ny_mils_summ),
                      ny_mins_summ = as.vector(ny_mins_summ)))

colnames(ny_summary_stats) <- c("Min", "1st Qu", "Median", "Mean",
                             "3rd Qu", "Max")

ph_mins_summ <- summary(play_PH$minutes)
ph_mils_summ <- summary(play_PH$miles)

ph_summary_stats <- t(data.frame(ph_mils_summ = as.vector(ph_mils_summ),
                      ph_mins_summ = as.vector(ph_mins_summ)))

colnames(ph_summary_stats) <- c("Min", "1st Qu", "Median", "Mean",
                             "3rd Qu", "Max")

dc_mins_summ <- summary(play_DC$minutes)
dc_mils_summ <- summary(play_DC$miles)

dc_summary_stats <- t(data.frame(dc_mils_summ = as.vector(dc_mils_summ),
                      dc_mins_summ = as.vector(dc_mins_summ)))

colnames(dc_summary_stats) <- c("Min", "1st Qu", "Median",
                                "Mean","3rd Qu", "Max")

acity_summs <- t(rbind(dc_summary_stats, ny_summary_stats, ph_summary_stats))
acity_summs1 <- as.data.frame(acity_summs)
acity_summs1 <-  acity_summs1[c(2, 6,4,1,5,3)]

colnames(acity_summs1) <- c("DC minutes","Philly minutes","NYC minutes", "DC miles", "Philly miles", "NYC miles")

##sds for all
x <- data.frame(
  q <- sd(play_DC$minutes),
  w <- sd(play_PH$minutes),
  e <- sd(play_NY$minutes),
  a <- sd(play_DC$miles),
  s <- sd(play_PH$miles),
  d <- sd(play_NY$miles))

colnames(x) <- c("DC minutes","Philly minutes","NYC minutes", "DC miles", "Philly miles", "NYC miles")

acity_summs1 <- rbind(acity_summs1, x)
rownames(acity_summs1) <-  c("Min", "1st Qu", "Median", "Mean",
                             "3rd Qu", "Max", "St. Dev.")

acity_summs1 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")
```

So let's start off by congratulating our nation's capital with running away with the competition. Everything about their summary statistics tells the game show contestents to hope to see `Washington DC` next to their assigned coordinates.

New York City and Philadelphia are actually in pretty close in everything related to time. Unfortunatly the average milage trip in NYC was 9 miles (and 12 in the 3rd quartile) longer than in Philly! So, if you're a real resident of NYC, you're getting some value for your time.

The only two concerning items are the Maxs and Standard Deviations' of New York and Philadelphia.

## Visual Exploration {-}

```{r scatterplots, echo = FALSE, fig.width=10}
DCscatter <- ggplot() +
  geom_point(data = play_DC, aes(x = miles, y = minutes),
                                     color = "#ff7400") +
  ggtitle("DC")+
  theme(legend.position="none")

PHscatter <- ggplot() +
  geom_point(data = play_PH, aes(x = miles, y = minutes),
                                     color = "#F14628") +
  ggtitle("Philly")+
  theme(legend.position="none")

NYscatter <- ggplot() +
  geom_point(data = play_NY, aes(x = miles, y = minutes),
                                     color = "#2360a5") +
  ggtitle("NYC")+
  theme(legend.position="none")

```

```{r histograms, echo = FALSE, message= FALSE, fig.width=10}
DChisto <- ggplot() +
  geom_histogram(data = play_DC, aes(x = minutes),fill = "#ff7400") +
  ggtitle("DC")+
  theme(legend.position="none")

PHhisto <- ggplot() +
  geom_histogram(data = play_PH, aes(x = minutes),fill = "#F14628") +
  ggtitle("Philly")+
  theme(legend.position="none")

NYhisto <- ggplot() +
  geom_histogram(data = play_NY, aes(x = minutes),fill = "#2360a5") +
  ggtitle("NYC")+
  theme(legend.position="none")

```

```{r , echo = FALSE, message= FALSE, fig.width=10, fig.height = 10}
grid.arrange(DChisto, PHhisto, NYhisto,DCscatter, PHscatter, NYscatter, ncol = 3)

```

```{r echo=FALSE, warning=FALSE}
df <- data.frame(
  DC = c(68.34, 61.05,7.29 ),
  Philly = c(142.7, 134.1, 8.6),
  NYC = c(468.5, 302.6, 165.8)
)

rownames(df) <- c("Total Area", "Land Area", "Water Area")

kable(df, "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right")
```

As suspected, we can see some Right Tail Extremes in the histograms of minutes for both NYC and Philadelphia. On the other hand, DC what a beauty. What causes this? My hypothesis is that it has to do with the size differences between the cities. You can see this in the table of Areas. This also, I think, helps explain the appearance of more`NULL` values in the Philadelphia and New York routes because of the much larger amount of water area within city limits.

Talk about skew here more. Maybe even a table.

## Evil Points {-}

In order to look at these outliers within the NYC and Philly dataset, I would like to put forward an axiom:

> Any trip over three hours, in this situation, can be defined as "Shit Luck", and any trip over five hours is straight up Evil.

Below I've split the histograms up to show these three different types of trips. normal trips (less than 3 hours) are in light blue, bad trips are in dark blue (3 - 5 hours), and truly evil points in black (5+ hours).

```{r scatter_badpoints_id, echo=FALSE, fig.width=10}

foo <- play_PH %>%
  mutate(comfort = ifelse(minutes >= 300, "evil",
              ifelse(minutes < 300 & minutes >= 181, "bad","normal"
                     )))

foo1 <- play_NY %>%
  mutate(comfort = ifelse(minutes >= 300, "evil",
              ifelse(minutes < 300 & minutes >= 181, "bad","normal"
                     )))

NYscatter2 <- ggplot() +
  geom_point(data = foo1, aes(x = miles, y = minutes, color = comfort))+
  scale_color_manual(values=c( "#0f16a8","black","#32a1fc"))+
  ggtitle("NYC")+
  theme(legend.position="none")

PHscatter2 <- ggplot() +
  geom_point(data = foo, aes(x = miles, y = minutes, color = comfort))+
  scale_color_manual(values=c( "#0f16a8","black","#32a1fc"))+
  ggtitle("PH")+
  theme(legend.position="none")

DCscatter2 <- ggplot() +
  geom_point(data = play_DC, aes(x = miles, y = minutes),
                                     color = "#32a1fc") +
  ggtitle("DC")+
  theme(legend.position="none")

grid.arrange(DCscatter2, PHscatter2, NYscatter2, ncol = 3)

```
<br>
<br>

```{r shit_trips,echo = FALSE,  warning = FALSE}
badluck_ny <- play_NY %>%
  filter(minutes >=180)

badluck_ph <- play_PH %>%
  filter(minutes >= 180)

badluck_dc <- play_DC %>%
  filter(minutes >= 180)
#Two hours
badluck_ny1 <- play_NY %>%
  filter(minutes >=120)

badluck_ph1 <- play_PH %>%
  filter(minutes >= 120)

badluck_dc1 <- play_DC %>%
  filter(minutes >= 120)

#percent of trips that are absurd
df <- data.frame(DC =
round((nrow(badluck_dc)/nrow(play_DC))*100, digits = 2),
PH = 
round((nrow(badluck_ph)/nrow(play_PH))*100, digits = 2),
NY = 
round((nrow(badluck_ny)/nrow(play_NY))*100, digits = 2))

df1 <- data.frame(DC =
round((nrow(badluck_dc1)/nrow(play_DC))*100, digits = 2),
PH = 
round((nrow(badluck_ph1)/nrow(play_PH))*100, digits = 2),
NY = 
round((nrow(badluck_ny1)/nrow(play_NY))*100, digits = 2))

df <- rbind(df1, df)
rownames(df) <- c( "2 hour trip", "3 hour trip")

```

###Percent of Bad Trips {-}
```{r shit_trips_table, echo=FALSE, warning=FALSE}
kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")

```

What we see here is that most of the bad points are not truly evil (only 8% and 6% respectively). So, what are those points? I've decided to restrict these points to being above 400 minutes (you'll see why in a second).

### NYC Evil Trips

```{r, echo = FALSE}

evil <- play_NY %>%
  filter(minutes >400)

#GGPLOT
foopath <- evil %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))

foopath1 <-evil %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))

foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along <- gg_ny +
  geom_point(data = foopath, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

###LEAFLET
leaf_ny <- 
leaflet(ny) %>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(lng = evil$startLon, lat = evil$startLat,
                  radius = 2, color = "red")
```

```{r echo = FALSE}
along
leaf_ny
```
<br>

So we can see that all `298` trips that are above are from a singular location. Furthermore, the mean of those trips is a staggering `r round(mean(evil$minutes), digits = 2)`.

### Philadelphia Evil Trips

```{r echo = FALSE}
evil <- play_PH %>%
  filter(minutes >400)

#GGPLOT
foopath <- evil %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <-evil %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along <- gg_ph +
  geom_point(data = foopath, aes(x = longitude, y = latitude, color = type), size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

# can see 2-3 bad bad points

evil_ph_5 <- data.frame(
  latitude = c(39.88880, 39.90100, 40.05352,39.88880 ,39.89117),
  longitude =c(-75.20997, -75.21456, -75.23903, -75.20997, -75.14090)
)

leaflet_ph <- 
leaflet(philly) %>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addTiles() %>%
  addCircleMarkers(lng = evil_ph_5$longitude, lat = evil_ph_5$latitude,
                   radius = 2, color = "blue")

```

```{r echo = FALSE}
along
leaflet_ph
```

```{r shit_trips_table_philly, echo=FALSE, warning=FALSE}
tab_evil <- aggregate(evil[,1], list(evil$startLat), length)%>%
  arrange(desc(x))

tab_evil1 <- aggregate(evil[,1], list(evil$endLat), length)%>%
  arrange(desc(x))

tab_evil <- cbind(tab_evil[1:50,], tab_evil1[1:50,])
colnames(tab_evil) <- c("Bad Origin", "count", "Bad Destination", "count")
tab_evil <- tab_evil[1:5,]

kable(tab_evil, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")

```
<br>
<br>
So, unlike NYC, Philadelphia has several locations that are either hard to get to or from. This is apparent from the table to the right.

<br>
## Results of Trips < 3 Hours {-}

Let's say, for no reason at all, that the gameshow guarantees that no trip will be more than 3 hours. How do the results change?

```{r reduced_scatter, echo = FALSE, fig.width=10}
goodluck_ny <- play_NY %>%
  filter(minutes <=250)

goodluck_ph <- play_PH %>%
  filter(minutes <= 250)

PHscatter1 <- ggplot() +
  geom_point(data = goodluck_ph, aes(x = miles, y = minutes),
                                     color = "#F14628") +
  ggtitle("Philly")+
  theme(legend.position="none")

NYscatter1 <- ggplot() +
  geom_point(data = goodluck_ny, aes(x = miles, y = minutes),
                                     color = "#2360a5") +
  ggtitle("NYC")+
  theme(legend.position="none")

PHhisto2 <- ggplot() +
  geom_histogram(data = goodluck_ph, aes(x = minutes),fill = "#F14628") +
  ggtitle("Philly")+
  theme(legend.position="none")

NYhisto2 <- ggplot() +
  geom_histogram(data = goodluck_ny, aes(x = minutes),fill = "#2360a5") +
  ggtitle("NYC")+
  theme(legend.position="none")

```

```{r , echo = FALSE, message= FALSE, fig.width=10, fig.height = 10}
grid.arrange(DChisto, PHhisto2, NYhisto2,DCscatter, PHscatter1, NYscatter1, ncol = 3)

```

```{r echo = FALSE}
acity_dat <- play_DC

foo <- acity_dat %>%
  arrange(minutes)
#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_short_dc <- gg_dc +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#######
#getting top 10% sample
foo <- acity_dat %>%
  arrange(desc(minutes))

#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_long_dc <- gg_dc +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#some of the middle

foo <- acity_dat %>%
  arrange(desc(minutes))
#getting 10% of rows
x <- (nrow(foo)*.1)*.5

x <- (0.5*nrow(acity_dat))-x
y <- (0.5*nrow(acity_dat))+x

foo <- foo[x:y,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_med_dc <- gg_dc +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

```

```{r echo = FALSE}

acity_dat <- play_PH

foo <- acity_dat %>%
  arrange(minutes)
#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_short_ph <- gg_ph +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#######
#getting top 10% sample
foo <- acity_dat %>%
  arrange(desc(minutes))

#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_long_ph <- gg_ph +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#some of the middle

foo <- acity_dat %>%
  arrange(desc(minutes))
#getting 10% of rows
x <- (nrow(foo)*.1)*.5

x <- (0.5*nrow(acity_dat))-x
y <- (0.5*nrow(acity_dat))+x

foo <- foo[x:y,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_med_ph <- gg_ph +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

```

```{r echo = FALSE}

acity_dat <- play_NY

foo <- acity_dat %>%
  arrange(minutes)
#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_short_ny <- gg_ny +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#######
#getting top 10% sample
foo <- acity_dat %>%
  arrange(desc(minutes))

#getting 10% of rows
x <- nrow(foo)*.1

foo <- foo[1:x,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_long_ny <- gg_ny +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

#some of the middle

foo <- acity_dat %>%
  arrange(desc(minutes))
#getting 10% of rows
x <- (nrow(foo)*.1)*.5

x <- (0.5*nrow(acity_dat))-x
y <- (0.5*nrow(acity_dat))+x

foo <- foo[x:y,]%>%
  sample_n(50)

foopath <- foo %>%
  select(latitude = endLat, longitude = endLon, trip_id) %>%
  mutate(type = rep("destination"))
foopath1 <- foo %>%
  select(latitude = startLat, longitude = startLon, trip_id) %>%
  mutate(type = rep("origin"))
foopath <- rbind(foopath, foopath1)
foopath$type <- as.factor(foopath$type) 
foopath$trip_id <- as.factor(foopath$trip_id)

along_med_ny <- gg_ny +
  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
  theme(legend.position="none")

```

### Sample of Short, Medium, and Long Trips {-}
I wanted to see what some of the different length trips looked like in each city. Below is series of graphs showing short, medium, and long trips. This was done by taking sample of 50 trips from the bottom, middle, and top 10% of trips for each city.

```{r echo = FALSE, fig.height = 10, fig.width = 12}
grid.arrange(along_short_dc,along_med_dc, along_long_dc,
             along_short_ph,along_med_ph, along_long_ph,
             along_short_ny,along_med_ny, along_long_ny, ncol = 3)
```

##Neighborhoods {-}

As a last question, I was wondering where within each city was the best origin. Below are the top ten origins.

```{r echo = FALSE}
##Philly

foo <- play_PH %>%
  group_by(startLat)%>%
  filter(n()>2)%>%
  summarise_all(mean) %>%
  arrange(minutes)

#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
head_neighborhoods_ph <- data.frame(
        latitude = foo$startLat[1:10],
        longitude =foo$startLon[1:10])

along <- gg_ph +
  geom_point(data = head_neighborhoods_ph, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)

leaf_ph_neigh <- 
  leaflet(philly) %>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(lng = head_neighborhoods_ph$longitude, lat = head_neighborhoods_ph$latitude,
                   radius = 2, color = "red")

leaf_ph_neigh

```

<br>

```{r echo = FALSE}
###DC

foo <- play_DC %>%
  group_by(startLat)%>%
  filter(n()>3 )%>%
  summarise_all(mean) %>%
  arrange(minutes)

#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
head_neighborhoods_dc <- data.frame(
  latitude = foo$startLat[1:10],
  longitude =foo$startLon[1:10])

along <- gg_dc +
  geom_point(data = head_neighborhoods_dc, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)

leaf_dc_neigh <- 
  leaflet(dc) %>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(lng = head_neighborhoods_dc$longitude, lat = head_neighborhoods_dc$latitude,
                   radius = 2, color = "red")

leaf_dc_neigh
```

<br>

```{r echo = FALSE}
###NY

foo <- play_NY %>%
  group_by(startLat)%>%
  filter(n()>20 )%>%
  summarise_all(mean) %>%
  arrange(minutes)

#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
head_neighborhoods_ny <- data.frame(
  latitude = foo$startLat[1:10],
  longitude =foo$startLon[1:10])

along <- gg_ny +
  geom_point(data = head_neighborhoods_ny, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)

leaf_ny_neigh <- 
  leaflet(ny) %>%
  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
              opacity = 0.5 ,color = "white")%>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  addCircleMarkers(lng = head_neighborhoods_ny$longitude, lat = head_neighborhoods_ny$latitude,
                   radius = 2, color = "red")

leaf_ny_neigh
```

# Endnotes:

**************
[^1]: One of only 3 positions in life that take shotgun seat in front of good friend. The other two being parents and grandparents.

[^2]: This made the trip much more annoying. In Philadelphia they had purchased a new caspar mattress during a 4th of July sale. My friend had done this forgetting he'd offered me a ride...the mattress and I were quite squished in back seat.

[^3]: The reason for this is purely personal. I had spent the best 10 years of my childhood around Philadelphia. The three of us in the car had attended University of Maryland, and so had used DC public transportation regularly over the last 4 years. Lastly, my friends were two weeks from moving to NYC to start their lives together.

[^4]: [Wikipedia ShapefileArticle](https://en.wikipedia.org/wiki/Shapefile)

[^5]: I hope it's clear enough from the code chunk, but I want to make it clear that the numextract function is not my own. I found it on: http://stla.github.io/stlapblog/posts/Numextract.html

[^6]: Spoilder alert: not for long.

[^7]: status codes starting 3xx or 4xx spell out trouble. Also, can use the httr function `status_code()` to get this piece of information easier.

[^8]: This includes some absurd areas like within rivers and oceans.
[^9]: I used the [HERE API](https://developer.here.com/)

### 2019-01-02 content/post/2019-01-02-transit.Rmd

@@
 date: '2019-01-02'
-slug: transit
-categories: []
-tags: []
+categories: ["R"]
+tags: ["R Markdown", "plot", "regression"]
+output: 
+  html_document:
+    self_contained: no
 ---
 
-Please please work
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
+                      results='show', cache=FALSE, autodep=FALSE)
+
+knitr::opts_knit$set(root.dir = 'data/transit/')
+```
+
+```{r, include = FALSE, message= FALSE, warning = FALSE}
+# include all packages here if nec.
+library(kableExtra)
+library(tigris)
+library(tidyverse)
+library(leaflet)
+library(sp)
+library(gridExtra)
+library(DT)
+library(data.table)
+library(rowr)
+
+```
+
+```{r shapefiles, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}
+philly <- tracts(state = "PA", county = "Philadelphia")
+dc <- tracts(state = "District of Columbia")
+ny <- tracts(state = "NY", 
+             county = c("New York County", "Bronx County",
+                        "Kings County", "Queens County", "Richmond County")  )
+
+```
+
+```{r dataload, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}
+
+load("GET_url_test.RData")
+load("content_url_test.RData")
+load("final_url_test.RData")
+load("json_url_test.RData")
+
+load("googleway_raw_data.RData")
+
+load("philly_300_points.RData")
+
+load("aplay_PH.RData")
+load("aplay_DC.RData")
+load("aplay_NY.RData")
+
+```
+
+```{r base_ggplots, echo = FALSE,  message= FALSE, warning = FALSE, error= FALSE, results = "hide"}
+gg_ph <- 
+  ggplot() +
+  geom_polygon(data = philly, aes(long, lat, group = group), 
+               fill = "#051633", color="#051633") +
+  coord_map() +
+  labs(x = NULL, y = NULL) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y=element_blank(),
+        panel.background = element_rect(fill = '#e8edf7'),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank())
+
+gg_ny <- 
+  ggplot() +
+  geom_polygon(data = ny, aes(long, lat, group = group), 
+               fill = "#051633", color="#051633") +
+  coord_map() +
+  labs(x = NULL, y = NULL) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y=element_blank(),
+        panel.background = element_rect(fill = '#e8edf7'),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank())
+
+gg_dc <- 
+  ggplot() +
+  geom_polygon(data = dc, aes(long, lat, group = group), 
+               fill = "#051633", color="#051633") +
+  coord_map() +
+  labs(x = NULL, y = NULL) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y=element_blank(),
+        panel.background = element_rect(fill = '#e8edf7'),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank())
+
+```
+
+# Part 1: The Idea  {-}
+
+I was recently in a car with one of my best friends and his new wife[^1] (also a good friend of mine).
+We were making our way back from our parents joint 4th of July BBQ in Philadelphia towards Washington DC where we had all attended University.
+These two friends of mine were coming back down to prepare for their fast approaching move to NYC [^2].
+
+Car talk moved towards talking about different public transit systems we've used both around the country and world.
+Being poor students and recent graduates and all moving in the near future, we are three of us particulary interested in our city's system.
+This was, of course, mostly complaining about the systems and comparing their negatives.
+This led me to an, I think, interesting question:
+
+If you were to be dropped in a city at a random point within that cities limit and given another random point to get to, again within the city limits, which city would this be the easiest in?
+
+I will here be trying to begin answering that question. In order to simplify the scope of this exploration, I am limiting myself to looking at New York City, Washington DC, and Philadelphia[^3].
+
+##Plan of Attack 1 {-}
+###Packages  {-}
+This analysis will be occuring entirely in R, so the first step is to load up some necessary packages to get started.
+
+###Let's Get Started {-}
+
+The first idea I had in solving this was to create a box around each city, get the longitude-latitude codes of the two corners, then get random points using `points <-  runif(num_points, lower_long/lat_barrier, upper_long/lat_barrier)`.
+
+As an example:
+```{r}
+x <- data.frame( x= runif(3, 1,50), y = runif(3, 51, 100))
+```
+```{r, echo=FALSE}
+x %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
+                position = "float_right")
+```
+
+Runif will return to us a specified set of random numbers (here `3`) from a uniform distribution. It is a similar function to RandBetween in Excel. You can see the results in the table to the right.
+
+So, I needed to get approximated bounded corners for the city. The way I decided to do this was to use googlemaps functionality of finding longitude and latitude of any point clicked.
+
+```{r echo=FALSE, out.width = '100%'}
+knitr::include_graphics("x1.png")
+
+```
+```{r echo=FALSE, out.width = '35%'}
+knitr::include_graphics("Untitled.png")
+```
+
+<br>
+
+So, after some slight maneuvering, we can use the two sets of coordinates to create this:
+```{r echo = FALSE}
+leaflet(philly) %>%
+    addProviderTiles(providers$Stamen.Toner) %>%
+  addRectangles(
+    lng1=-75.381713, lat1=40.143756,
+    lng2=-74.949813, lat2=39.885019, fillColor  = "transparent"
+  )%>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")
+
+```
+
+Great!
+Okay, so we can use our coordinate codes of `-75.381713, 40.143756` and `-74.949813, 39.885019` within our `runif` function.
+As a test, I'll generate 100 random points and map them out.
+Below I've provided a sample of the points in a table.
+
+```{r}
+longitude <-runif(100, -75.381713, -74.949813)
+latitude <- runif(100, 39.885019, 40.143756)
+my_df <- data.frame(longitude, latitude)
+```
+
+```{r, echo=FALSE}
+my_df[1:10,] %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped"), full_width = FALSE,
+                position = "float_right")
+```
+
+```{r echo = FALSE}
+
+leaflet(philly) %>%
+  addProviderTiles(providers$Stamen.Toner) %>%
+  addRectangles(
+    lng1=-75.381713, lat1=40.143756,
+    lng2=-74.949813, lat2=39.885019, fillColor  = "transparent"
+  )%>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addCircleMarkers(data=my_df,radius =1.5,
+                   color= "#ff0505")
+
+```
+
+## A Clear Issue {-}
+This is obviously an issue.  
+Cities are not specificially oriented in the plot of the Earth's coordinate system, they are a "random" human choice.
+Even worse (becuase with motivation, the perspective of the graph can be changed), cities are weird shapes.
+Humans shaped these amazing metropolises that span whatever land they decide.
+A very cool thought, but highly inconvenient for me.
+
+# Part 2: Shapefiles and Progress {-}
+
+```{r echo=FALSE, out.width = '50%'}
+knitr::include_graphics("Septa_logo.png")
+
+```
+
+The main lesson from part 1 is that I'm going to have to be a little less lazy in this project. The minitua is that I needed:
+
+   1. a way to graph only the city limits itself  
+   2. a way to generate random points within that irregular polygon
+
+Googling told me that what I needed was a shapefile, "The shapefile format is a popular geospatial vector data format for geographic information system (GIS) software...The shapefile format can spatially describe vector features: points, lines, and polygons, representing, for example, water wells, rivers, and lakes."[^4]
+
+At first, it looked like this required me to download a bunch of files from the internet and load them all into R.
+At first because of the blessed package Tigris which allows users to download shapefiles from the US Census Bureau.
+
+The below code is using the TIGRIS package to grab the files for Philadelphia, New York City, and Washington D.C.
+```{r echo = FALSE, message= FALSE, warning = FALSE, fig.width=10, fig.height=5}
+
+grid.arrange(gg_ph, gg_ny, gg_dc, ncol = 3)
+
+```
+Okay great - we have our city polygons.
+## Getting Random Points {-}
+To work with the shapefiles in R, the package `sp` was used.
+This package has a great function `spsample` which magically spits out the random numbers.
+At first, I still thought I'd have to figure out how to divide the cities into shapes I could find random points within then recombine.
+
+I thought a good starting point would be to get 300 points, 150 origin and 150 destination, and find the directions between them. I'll show an example of that workflow using only Philadelphia.
+
+```{r, cache = TRUE}
+number_sample <- 150  #This is the number of trips we want to test
+city_file <- philly #This is what city to get points for
+
+origin <- spsample(city_file, n = number_sample, "random") %>%
+  as.data.frame()
+origin <- origin %>%
+  select(startLat = y , startLon = x) %>%
+  mutate(trip_id = 1:nrow(origin))
+
+destination <- spsample(city_file, n = number_sample, "random") %>%
+  as.data.frame()
+destination <- destination %>%
+  select(endLat = y , endLon = x) %>%
+  mutate(trip_id = 1:nrow(destination))
+
+#To combine the origins and destinations
+trip <- left_join(origin,destination,by = "trip_id") %>%
+  select(trip_id,startLat, startLon, endLat, endLon)
+```
+
+```{r, echo=FALSE,warning = FALSE}
+trip1 <- trip %>%
+  select("start latitude" = startLat, "start longitude" = startLon,  "end latitude" = endLat, "end longitude" = endLon )
+
+datatable(trip1, options = list(
+  dom = 't'
+))
+```
+<br>
+
+```{r, echo = FALSE}
+origin <- origin %>%
+  select(latitude = startLat, longitude = startLon, trip_id)%>%
+  mutate(type = rep("origin"))
+destination <- destination %>%
+  select(latitude = endLat, longitude = endLon, trip_id)%>%
+  mutate(type = rep("destination"))
+
+loc_both <- rbind(origin, destination)
+
+city_origin_destination <- gg_ph + 
+  geom_point(data = loc_both, aes(longitude, latitude, group = type, color =type), size =  1.5)+
+  scale_color_manual(values=c("#db7918", "#1dbdf7"))
+
+city_origin_destination  
+```
+
+Now if we want to view the "routes" themselves if the one went as the bird flew.
+
+```{r, echo = FALSE}
+
+trip_a2b_lines <- city_origin_destination +
+  geom_path(data=loc_both, aes(x=longitude, y=latitude, group=trip_id), size=0.5, alpha=0.5, color = "white")
+
+trip_a2b_lines
+```
+
+##Getting Routes {-}
+
+So now we have this shit.
+great package
+google way // attaches to google api
+
+```{r, echo = FALSE}
+#loading a predone dataframe so that the document doesn't
+#ping the API everytime I reknit...or ever really
+
+```
+
+```{r, eval = FALSE}
+googleway_raw_data <- lapply(1:nrow(trip), function(x){
+
+  google_distance(origin = (locations[x, 2:3]),
+                  destination = (locations[x, 4:5]),
+                  key = key,
+                  mode = "transit")
+
+})
+
+```
+This results in a list of lists where each holds the data of 1 trip.
+```{r}
+googleway_raw_data[[5]] #looking at the fifth list, which is also the fifth set of coordinates
+```
+
+So, some serious data munging is needed... *cue the montage music*
+
+```{r, warning= FALSE}
+foo <- do.call(what = "rbind",
+               args = lapply(googleway_raw_data, as.data.frame))
+
+foo <-  separate(foo, elements, into = c("km", "value", "duration", 
+                    "value2", "currency", "cost", "uk", "status2"), ",")
+
+foo <- 
+  select(foo, destination_addresses, origin_addresses, km, duration, cost, status)
+head(foo)
+```
+
+Now to get rid of all that annoying fluff[^5]
+
+```{r}
+##source of numextract function: http://stla.github.io/stlapblog/posts/Numextract.html
+numextract <- function(string){ 
+  str_extract(string, "\\-*\\d+\\.*\\d*")
+}
+
+foo$km <- lapply(foo$km, numextract)%>%
+  as.numeric()
+foo$cost <- substring(foo$cost, 11)
+foo$cost <- substr(foo$cost,1,nchar(foo$cost)-1) %>%
+  as.numeric()
+foo$duration <- substring(foo$duration, 26)
+foo$duration <- substr(foo$duration,1,nchar(foo$duration)-1)
+
+#Splitting up duration from XX hour(s) XX minute(s), to seperate columns
+
+foo1 <- separate(foo, duration, into = c("hours", "minutes"), sep = -7)
+foo1$hours <- gsub("[^0-9]", "", foo1$hours)
+foo1$minutes <- gsub("[^0-9]", "",foo1$minutes)
+foo1$hours <- as.numeric(foo1$hours)
+foo1$minutes <- as.numeric(foo1$minutes)
+
+#for trips that take less than an hour
+#doesnt murk with invalid trips because of other apparent NAS
+foo1$hours[is.na(foo1$hours)] <- 0
+
+#creating a total minutes section
+foo1 <- foo1 %>% 
+  mutate(duration_minutes = (hours*60)+minutes)%>%
+  mutate(trip_id = 1:nrow(foo1))
+
+#merging with trip to have coordinates
+full_AB_GOOGLE <- left_join(foo1, trip,by = "trip_id")
+
+```
+
+```{r, echo=FALSE,warning = FALSE}
+full_AB_GOOGLE1 <- full_AB_GOOGLE
+
+full_AB_GOOGLE1$startLat <- round(full_AB_GOOGLE1$startLat, digits = 4)
+full_AB_GOOGLE1$startLon <- round(full_AB_GOOGLE1$startLon, digits = 4)
+full_AB_GOOGLE1$endLat <- round(full_AB_GOOGLE1$endLat, digits = 4)
+full_AB_GOOGLE1$endLon <- round(full_AB_GOOGLE1$endLon, digits = 4)
+full_AB_GOOGLE1 <- full_AB_GOOGLE1 %>%
+  select(origin_addresses, destination_addresses, duration_minutes, km, cost, startLat, startLon, endLat, endLon)
+
+datatable(full_AB_GOOGLE1, options = list(
+  dom = 'ltp'
+))
+```
+
+## Our first "good" data {-}
+Let's take a quick peak at what the data tells us before moving on.
+
+```{r warning = FALSE}
+#selecting a few choice variables and adding miles
+foo <- full_AB_GOOGLE %>%
+  select(trip_id, kilometers = km, duration = duration_minutes, cost)%>%
+  mutate(miles = round(kilometers/1.609344, digits =1))
+
+miles <- as.data.frame(c(summary(foo$miles)))
+duration <-as.data.frame(c(summary(foo$duration)))
+summary1 <- cbind(miles, duration)%>%
+  round(digits =2)
+colnames(summary1) <- c("miles", "durations")
+
+summary1 %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
+                position = "float_right")
+
+```
+
+```{r echo = FALSE, message= FALSE, warning = FALSE}
+
+histx <- 
+ggplot(data = foo, aes(duration)) + 
+  geom_histogram(bins = 10, fill = "#1f4fa3")+
+  ggtitle("Histogram of Trip Duration")
+
+scatter <- 
+ggplot() +
+  geom_point(data =foo, aes(x = duration, y = miles), color = "#f14728")+
+  ggtitle("Scatterplot of Duration vs. Miles")
+
+grid.arrange(histx, scatter,  ncol = 2)
+
+```
+
+# Part 3: More Trips {-}
+
+```{r echo=FALSE, out.width = '25%'}
+knitr::include_graphics("mta_logo.png")
+
+```
+
+At this point we've found directions from 150 points to a seperate set of 150 points that were derived from within each city's shapefile. 
+This, honestly, seemed a bit weak to me. It answers the question...kind of. The hypothetical situation seemed to call for, well, *more points*.
+
+So the idea is this:
+
+   1. Forget about origins and destinations and combine the 300 points into one list  
+   2. Find directions from each point to the other 299 points
+
+```{r warning = FALSE}
+#Changing the column names to allow binding
+col_names <- c("latitude", "longitude" )
+
+#Getting the columns wanted from origin and destination, then changing the column names
+coordinates <- list(origin, destination)%>%
+  lapply( function(x) {
+  subset(x, select= -trip_id)
+  }) %>%
+  lapply(setNames, nm = col_names)
+
+#combining the origin and destination coordinate points
+loc_both <- rbind(coordinates[[1]],coordinates[[2]])
+
+#splitting each individual set of coordinates into their own dataframe
+a <- loc_both %>%
+  apply(1, as.data.frame)
+
+#Creating a function that takes each dataframe of a list(here a), replicates the rows the number of times needed, then creates another 2 columns where the other 300 points go (including a duplicate)
+multi.lapply <- function(a){
+  len <- nrow(loc_both)
+  a <- lapply(a, function(x) apply(x, 1, t))
+  c <- lapply(a, as.data.frame)
+  d <- lapply(c, t)
+  e <- lapply(d, function(x)  apply(x, 2, rep, len))
+  f <- lapply(e, function(x) cbind(x, 2, loc_both))
+  new_col_name <-  c("startLat", "startLon", "trip", "endLat", "endLon")
+  g <- lapply(f, setNames, nm = new_col_name)
+  h <- bind_rows(g)
+  return(h)
+}
+
+#Running the created function
+philly_300_points <- multi.lapply(a)
+philly_300_pointsraw <-philly_300_points[c(3,1,2,4,5)]
+
+#removing the duplicate rows
+philly_300_points <- philly_300_pointsraw %>%
+  filter(startLat != endLat, startLon != endLon)
+
+head(philly_300_points)
+```
+
+So now we have a dataframe of `89700` rows.
+Which is a subset of the `300*300 = 90,000` observations we expected.
+This is because each trip has one duplicate row (where the starting longitude = the ending longitude and the starting latitude = the ending latitude).
+
+##Getting the Routes using Google API {-}
+As before, we'll be using the googleway package to access the google maps API [^6]. However, I had no interest in letting this function run for the next however much time.
+
+A friend of mine recommended looking into parallel computing over the cloud.
+It was the perfect recommendation.
+I setup an AWS account and got an Rstudio environment running, which was a bit of a headache.
+On July 18, I had finally gotten it up and running.
+That night I went to sleep and left my computer on as my data was processed.
+
+```{r eval = FALSE}
+
+pls <-  mclapply(1:nrow(tfoo), function(x){
+
+  google_distance(origin = (tfoo[x, 1:2]),
+                  destination = (tfoo[x, 4:5]),
+                  key = ky,
+                  mode = "transit")
+
+})
+```
+
+#Part 4: A Catastrophic Issue {-}
+
+```{r echo=FALSE, out.width = '25%'}
+knitr::include_graphics("wmata_logo.png")
+
+```
+
+Okay, I woke up the morning of the July 19th and found the two biggest roadblocks of the project waiting at my computer.
+What made it castrophic was that, along with these two issues, I had my first set of 90,000 routes.
+
+The first issue was that that day, Google had begun charging for their previously free API.
+A disastear for my student pockets, and so close to the solution.
+Well, I was too close to not try and find a replacement.
+
+The second issue was that my use of AWS had cost around $50, not something I was expected. Ooops.
+
+Well let's try and solve the first issue first.
+
+##Finding a Replacement for Google Routes API {-}
+Finding another API was only a mild headache.
+The most annoying aspect was finding one that suppored transit routes and not exclusively driving directions.
+To the rescue: Here API.
+It had everything I wanted, except that no kind soul had made an R package to easily access it.
+It became pretty obvious during that day of research that Here was my best option.
+I would have to write some of my own functions to get it to work.
+
+###Functions {-}
+Basically, I want to take my 3 lists of 90,000 rows and 360,000 indivudal cooridnates and feed them into the Here API.
+
+The API requires that the link I use look like:
+
+> "route.cit.api.here .com/routing/7.2/calculateroute.json?waypoint0=52.5208%2C13.4093&waypoint1=52.5034%2C13.3295&mode=fastest%3BpublicTransport &combineChange=true&app_id=DemoAppId01082013GAL&app_code=AJKnXv84fjrb0KIHawS0Tg"
+
+All coordinates needed to be adjusted to be strings that resemble `52.5208%2C13.4093`.
+
+```{r}
+here_mcoord_fix <- function(df){
+  latitude <- df[,1]
+  longitude <- df[,2]
+  url2 <- paste0(longitude, "%2C", latitude)
+  return(url2)
+}
+```
+
+Now, there are several constants in the URL that can be set to variables. Then everything can be easily pasted together into proper URLS.
+
+```{r}
+base_url <- "https://route.cit.api.here.com/routing/7.2/calculateroute.json"
+xmode <- "&mode=fastest%3BpublicTransport&combineChange=true"
+#these two are different per API User
+id <- "&app_id=SOME_NUMBERS_AND_LETTERS&"
+code <- "app_code=SOME_NUMBERS_AND_LETTERS&departure=2018-07-18T11:00:00-05:00"
+```
+
+And here's the function that pulls everything together. It contains the `here_mcoord_fix` function within to cut steps down.
+
+```{r}
+get_here_urls <- function(origin, destination){
+  xorigin <- here_mcoord_fix(origin)
+  xdestination <- here_mcoord_fix(destination)
+  z <- paste0(base_url,"?waypoint0=",xorigin, "&waypoint1=",
+              xdestination, xmode, id, code, sep="")
+  return(z) 
+}
+
+```
+
+So let me show the output:
+```{r}
+origin <- philly_300_points%>%
+  select(startLon, startLat)
+destination <- philly_300_points%>%
+  select(endLon, endLat)
+
+urls <- get_here_urls(origin, destination)
+
+#Let's take a peak at a random 1 of 89700 URLs we've created.
+urls[runif(1, 1,89700)]
+
+```
+
+##Running this Code without Emptying my Pockets {-}
+So, I decided not to use AWS anymore...maybe if I had deep pockets I would, but I had decided to find an alternative.
+The first thing I learned was that I did not[^1] really fully understand parallel computing.
+There were better functions than `mclapply` available.
+The second thing I realized that I was obsessed with doing the work on the cloud.
+On one hand, the cloud is definietly better.
+On the other hand, only if you've got that cash-money.
+
+The solution was to use the `DoParallel` package with computers that have multiple cores, thus performing parallel computations on each core.
+
+###Getting DoParallel Up and Running {-}
+```{r, eval = FALSE}
+#Setting up parallel with one less core than available on the computer. This is to avoid intense crashing.
+no_cores <- detectCores() - 1  
+cl <- makeCluster(no_cores)  
+registerDoParallel(cl)
+
+#To see if that worked: if it returns 1 - then it didn't work
+getDoParWorkers()
+
+#To end the parallel cores
+registerDoSEQ()
+getDoParWorkers()
+
+```
+
+###Getting the functions ready to access the here API {-}
+The first step is getting it running before going parallel and seeing what the outputs look like.
+
+There are some messy looking results so I've tabbed this section to maintain readability. Feel free to look through it if you're interested. Each tab is a view of each output.
+
+#### Everything {-}
+The first step is getting it running before going parallel and seeing what the outputs look like:
+
+```{r eval = FALSE}
+
+#We'll use one random url to test
+url_test <- urls[runif(1, 1,89700)]
+require(httr)
+require(jsonlite)
+
+GET_url_test <- GET(url_test)
+content_url_test <-  content(GET_url_test, "text")
+json_url_test <-   fromJSON(content_url_test, flatten = TRUE)
+final_url_test <- as.data.frame(json_url_test)
+
+```
+
+#### Step 1 {-}
+```{r}
+GET_url_test
+```
+There are a few good pieces of news in this code:
+
+  *The status is 200 meaning it worked[^7]
+  *The data is in JSON format
+  *The size is not absurd
+
+#### Step 2 {-}
+```{r eval = FALSE}
+content_url_test <-  content(GET_url_test, "text")
+```
+
+```{r echo = FALSE}
+content_url_test
+```
+
+#### Step 3 {-}
+
+```{r eval = FALSE}
+json_url_test <-   fromJSON(content_url_test, flatten = TRUE)
+
+```
+
+```{r echo = FALSE}
+json_url_test
+```
+
+#### step 4 {-}
+```{r eval = FALSE}
+ final_url_test <- as.data.frame(json_url_test)
+```
+
+```{r echo = FALSE}
+final_url_test
+
+```
+
+###Preparing Looping Functions for Parallel {-}
+
+```{r eval = FALSE}
+#gets the Data from HERE API
+
+get_city = foreach(i=urls, .packages='httr') %dopar% {
+    GET(i)
+  }
+
+status_code(results[[1]]) #Can check some to ensure it worked
+
+#Gets the Json Content
+content_city = foreach(i=get_city, .packages='httr') %dopar% {
+  content(i, "text")
+}
+
+#munging
+#makes the json data semi pretty
+json_city = foreach(i=content_city, .packages='jsonlite') %dopar% {
+  fromJSON(i, flatten = TRUE)
+}
+
+df_city = foreach(i=json_city) %dopar% {
+    as.data.frame(i)
+  }
+
+full <- rbindlist(df_city, fill = TRUE)
+
+```
+
+###Plan of Attack {-}
+To solve my 2nd large issue, I decided to leverage the resources of my university.
+Generally, the idea is to split up the urls into multiple lists, then run each set of URLs independently on several computers.
+
+To split up the URLs:
+```{r}
+list_of_URLs <- split(urls, ceiling(seq_along(urls)/15000))
+str(list_of_URLs)
+
+```
+
+Then on each computer I would run the code, but first specify:
+```{r}
+urls<- list_of_URLs[[2]] #insert whatever subset you want to use
+str(urls)
+
+```
+
+This worked magically.
+I used several computers over 3 days and got all the data I need.
+The internet and books constantly preached that 80% of these tasks are data collection / wrangling and only 20% was the actual analysis.
+I think I finally understand what they mean...But now for the fun!
+The conclusions and results!!!
+
+# Part 5: Results and Conclusion {-}
+
+## Summary of Results {-}
+
+```{r echo = FALSE, warning=FALSE}
+fooph <- play_PH %>%
+  group_by(startLat)%>%
+  filter(n()<2)%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+fooDC <- play_DC %>%
+  group_by(startLat)%>%
+  filter(n()<3 )%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+foony <- play_NY %>%
+  group_by(startLat)%>%
+  filter(n()<20 )%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+acity_row <- data.frame(
+DC = c(nrow(play_DC), 90000-nrow(play_DC), nrow(fooDC), 100*((nrow(play_DC)+nrow(fooDC))/90000)),
+Philly = c(nrow(play_PH),90000 -nrow(play_PH),nrow(fooph), 100*((nrow(play_DC) + nrow(fooph))/90000)),
+NYC = c(nrow(play_NY),90000 -nrow(play_NY),nrow(foony),  100*((nrow(play_DC) + nrow(foony))/90000))
+)
+
+rownames(acity_row) <- c("Rows Remaining", "Rows NULL","Got through","Total Percent Removed")
+
+x <- as.matrix(acity_row)
+
+```
+
+```{r echo=FALSE, warning=FALSE}
+kable(x, "html") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right")
+```
+
+Before diving into this crazy results chart, I need to admit something...I don't know what I'm talking about when it comes to [GIS](https://en.wikipedia.org/wiki/Geographic_information_system).
+If I were a real maps whiz, I would have limited my city polygons to only include areas people lived.
+There were issues with several areas within each city.
+
+  1. Many of these issue areas resulted in, mostly, `NULL` responses from the API I used to get direction routes [^8]. These were easily removed. 
+  2.
+The majority of the rest were a little more *slippery* and got results from the API.
+An example of one of these were a point in the river would pull `298 NULLS` but would get directions to another random point within the river.
+These were also easily removed (except for a small few that will be discussed below).
+  3.
+The last bit are a bit more of a question.
+A question of how evil these made-up gameshow creators are.
+For example, several points fell into points deep within large philadelphian parks and began with 20-45 minutes of walking.
+These were kept in[^9].
+
+```{r summary_stats, echo = FALSE, warning = FALSE}
+ny_mins_summ <- summary(play_NY$minutes)
+ny_mils_summ <- summary(play_NY$miles)
+
+ny_summary_stats <- t(data.frame(ny_mils_summ = as.vector(ny_mils_summ),
+                      ny_mins_summ = as.vector(ny_mins_summ)))
+
+colnames(ny_summary_stats) <- c("Min", "1st Qu", "Median", "Mean",
+                             "3rd Qu", "Max")
+
+ph_mins_summ <- summary(play_PH$minutes)
+ph_mils_summ <- summary(play_PH$miles)
+
+ph_summary_stats <- t(data.frame(ph_mils_summ = as.vector(ph_mils_summ),
+                      ph_mins_summ = as.vector(ph_mins_summ)))
+
+colnames(ph_summary_stats) <- c("Min", "1st Qu", "Median", "Mean",
+                             "3rd Qu", "Max")
+
+dc_mins_summ <- summary(play_DC$minutes)
+dc_mils_summ <- summary(play_DC$miles)
+
+dc_summary_stats <- t(data.frame(dc_mils_summ = as.vector(dc_mils_summ),
+                      dc_mins_summ = as.vector(dc_mins_summ)))
+
+colnames(dc_summary_stats) <- c("Min", "1st Qu", "Median",
+                                "Mean","3rd Qu", "Max")
+
+acity_summs <- t(rbind(dc_summary_stats, ny_summary_stats, ph_summary_stats))
+acity_summs1 <- as.data.frame(acity_summs)
+acity_summs1 <-  acity_summs1[c(2, 6,4,1,5,3)]
+
+colnames(acity_summs1) <- c("DC minutes","Philly minutes","NYC minutes", "DC miles", "Philly miles", "NYC miles")
+
+##sds for all
+x <- data.frame(
+  q <- sd(play_DC$minutes),
+  w <- sd(play_PH$minutes),
+  e <- sd(play_NY$minutes),
+  a <- sd(play_DC$miles),
+  s <- sd(play_PH$miles),
+  d <- sd(play_NY$miles))
+
+colnames(x) <- c("DC minutes","Philly minutes","NYC minutes", "DC miles", "Philly miles", "NYC miles")
+
+acity_summs1 <- rbind(acity_summs1, x)
+rownames(acity_summs1) <-  c("Min", "1st Qu", "Median", "Mean",
+                             "3rd Qu", "Max", "St. Dev.")
+
+acity_summs1 %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")
+```
+
+So let's start off by congratulating our nation's capital with running away with the competition.
+Everything about their summary statistics tells the game show contestents to hope to see `Washington DC` next to their assigned coordinates.
+
+New York City and Philadelphia are actually in pretty close in everything related to time.
+Unfortunatly the average milage trip in NYC was 9 miles (and 12 in the 3rd quartile) longer than in Philly!
+So, if you're a real resident of NYC, you're getting some value for your time.
+
+The only two concerning items are the Maxs and Standard Deviations' of New York and Philadelphia.
+
+## Visual Exploration {-}
+
+```{r scatterplots, echo = FALSE, fig.width=10}
+DCscatter <- ggplot() +
+  geom_point(data = play_DC, aes(x = miles, y = minutes),
+                                     color = "#ff7400") +
+  ggtitle("DC")+
+  theme(legend.position="none")
+
+PHscatter <- ggplot() +
+  geom_point(data = play_PH, aes(x = miles, y = minutes),
+                                     color = "#F14628") +
+  ggtitle("Philly")+
+  theme(legend.position="none")
+
+NYscatter <- ggplot() +
+  geom_point(data = play_NY, aes(x = miles, y = minutes),
+                                     color = "#2360a5") +
+  ggtitle("NYC")+
+  theme(legend.position="none")
+
+```
+
+```{r histograms, echo = FALSE, message= FALSE, fig.width=10}
+DChisto <- ggplot() +
+  geom_histogram(data = play_DC, aes(x = minutes),fill = "#ff7400") +
+  ggtitle("DC")+
+  theme(legend.position="none")
+
+PHhisto <- ggplot() +
+  geom_histogram(data = play_PH, aes(x = minutes),fill = "#F14628") +
+  ggtitle("Philly")+
+  theme(legend.position="none")
+
+NYhisto <- ggplot() +
+  geom_histogram(data = play_NY, aes(x = minutes),fill = "#2360a5") +
+  ggtitle("NYC")+
+  theme(legend.position="none")
+
+```
+
+```{r , echo = FALSE, message= FALSE, fig.width=10, fig.height = 10}
+grid.arrange(DChisto, PHhisto, NYhisto,DCscatter, PHscatter, NYscatter, ncol = 3)
+
+```
+
+```{r echo=FALSE, warning=FALSE}
+df <- data.frame(
+  DC = c(68.34, 61.05,7.29 ),
+  Philly = c(142.7, 134.1, 8.6),
+  NYC = c(468.5, 302.6, 165.8)
+)
+
+rownames(df) <- c("Total Area", "Land Area", "Water Area")
+
+kable(df, "html") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right")
+```
+
+As suspected, we can see some Right Tail Extremes in the histograms of minutes for both NYC and Philadelphia.
+On the other hand, DC what a beauty.
+What causes this?
+My hypothesis is that it has to do with the size differences between the cities.
+You can see this in the table of Areas.
+This also, I think, helps explain the appearance of more`NULL` values in the Philadelphia and New York routes because of the much larger amount of water area within city limits.
+
+Talk about skew here more. Maybe even a table.
+
+## Evil Points {-}
+
+In order to look at these outliers within the NYC and Philly dataset, I would like to put forward an axiom:
+
+> Any trip over three hours, in this situation, can be defined as "Shit Luck", and any trip over five hours is straight up Evil.
+
+Below I've split the histograms up to show these three different types of trips.
+normal trips (less than 3 hours) are in light blue, bad trips are in dark blue (3 - 5 hours), and truly evil points in black (5+ hours).
+
+```{r scatter_badpoints_id, echo=FALSE, fig.width=10}
+
+foo <- play_PH %>%
+  mutate(comfort = ifelse(minutes >= 300, "evil",
+              ifelse(minutes < 300 & minutes >= 181, "bad","normal"
+                     )))
+
+foo1 <- play_NY %>%
+  mutate(comfort = ifelse(minutes >= 300, "evil",
+              ifelse(minutes < 300 & minutes >= 181, "bad","normal"
+                     )))
+
+NYscatter2 <- ggplot() +
+  geom_point(data = foo1, aes(x = miles, y = minutes, color = comfort))+
+  scale_color_manual(values=c( "#0f16a8","black","#32a1fc"))+
+  ggtitle("NYC")+
+  theme(legend.position="none")
+
+PHscatter2 <- ggplot() +
+  geom_point(data = foo, aes(x = miles, y = minutes, color = comfort))+
+  scale_color_manual(values=c( "#0f16a8","black","#32a1fc"))+
+  ggtitle("PH")+
+  theme(legend.position="none")
+
+DCscatter2 <- ggplot() +
+  geom_point(data = play_DC, aes(x = miles, y = minutes),
+                                     color = "#32a1fc") +
+  ggtitle("DC")+
+  theme(legend.position="none")
+
+grid.arrange(DCscatter2, PHscatter2, NYscatter2, ncol = 3)
+
+```
+<br>
+<br>
+
+```{r shit_trips,echo = FALSE,  warning = FALSE}
+badluck_ny <- play_NY %>%
+  filter(minutes >=180)
+
+badluck_ph <- play_PH %>%
+  filter(minutes >= 180)
+
+badluck_dc <- play_DC %>%
+  filter(minutes >= 180)
+#Two hours
+badluck_ny1 <- play_NY %>%
+  filter(minutes >=120)
+
+badluck_ph1 <- play_PH %>%
+  filter(minutes >= 120)
+
+badluck_dc1 <- play_DC %>%
+  filter(minutes >= 120)
+
+#percent of trips that are absurd
+df <- data.frame(DC =
+round((nrow(badluck_dc)/nrow(play_DC))*100, digits = 2),
+PH = 
+round((nrow(badluck_ph)/nrow(play_PH))*100, digits = 2),
+NY = 
+round((nrow(badluck_ny)/nrow(play_NY))*100, digits = 2))
+
+df1 <- data.frame(DC =
+round((nrow(badluck_dc1)/nrow(play_DC))*100, digits = 2),
+PH = 
+round((nrow(badluck_ph1)/nrow(play_PH))*100, digits = 2),
+NY = 
+round((nrow(badluck_ny1)/nrow(play_NY))*100, digits = 2))
+
+df <- rbind(df1, df)
+rownames(df) <- c( "2 hour trip", "3 hour trip")
+
+```
+
+###Percent of Bad Trips {-}
+```{r shit_trips_table, echo=FALSE, warning=FALSE}
+kable(df, "html") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")
+
+```
+
+What we see here is that most of the bad points are not truly evil (only 8% and 6% respectively).
+So, what are those points?
+I've decided to restrict these points to being above 400 minutes (you'll see why in a second).
+
+### NYC Evil Trips
+
+```{r, echo = FALSE}
+
+evil <- play_NY %>%
+  filter(minutes >400)
+
+#GGPLOT
+foopath <- evil %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+
+foopath1 <-evil %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along <- gg_ny +
+  geom_point(data = foopath, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+###LEAFLET
+leaf_ny <- 
+leaflet(ny) %>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addProviderTiles("Esri.WorldStreetMap") %>%
+  addCircleMarkers(lng = evil$startLon, lat = evil$startLat,
+                  radius = 2, color = "red")
+```
+
+```{r echo = FALSE}
+along
+leaf_ny
+```
+<br>
+
+So we can see that all `298` trips that are above are from a singular location. Furthermore, the mean of those trips is a staggering `r round(mean(evil$minutes), digits = 2)`.
+
+### Philadelphia Evil Trips
+
+```{r echo = FALSE}
+evil <- play_PH %>%
+  filter(minutes >400)
+
+#GGPLOT
+foopath <- evil %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <-evil %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along <- gg_ph +
+  geom_point(data = foopath, aes(x = longitude, y = latitude, color = type), size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+# can see 2-3 bad bad points
+
+evil_ph_5 <- data.frame(
+  latitude = c(39.88880, 39.90100, 40.05352,39.88880 ,39.89117),
+  longitude =c(-75.20997, -75.21456, -75.23903, -75.20997, -75.14090)
+)
+
+leaflet_ph <- 
+leaflet(philly) %>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addTiles() %>%
+  addCircleMarkers(lng = evil_ph_5$longitude, lat = evil_ph_5$latitude,
+                   radius = 2, color = "blue")
+
+```
+
+```{r echo = FALSE}
+along
+leaflet_ph
+```
+
+```{r shit_trips_table_philly, echo=FALSE, warning=FALSE}
+tab_evil <- aggregate(evil[,1], list(evil$startLat), length)%>%
+  arrange(desc(x))
+
+tab_evil1 <- aggregate(evil[,1], list(evil$endLat), length)%>%
+  arrange(desc(x))
+
+tab_evil <- cbind(tab_evil[1:50,], tab_evil1[1:50,])
+colnames(tab_evil) <- c("Bad Origin", "count", "Bad Destination", "count")
+tab_evil <- tab_evil[1:5,]
+
+kable(tab_evil, "html") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "float_right")
+
+```
+<br>
+<br>
+So, unlike NYC, Philadelphia has several locations that are either hard to get to or from. This is apparent from the table to the right.
+
+<br>
+## Results of Trips < 3 Hours {-}
+
+Let's say, for no reason at all, that the gameshow guarantees that no trip will be more than 3 hours. How do the results change?
+
+```{r reduced_scatter, echo = FALSE, fig.width=10}
+goodluck_ny <- play_NY %>%
+  filter(minutes <=250)
+
+goodluck_ph <- play_PH %>%
+  filter(minutes <= 250)
+
+PHscatter1 <- ggplot() +
+  geom_point(data = goodluck_ph, aes(x = miles, y = minutes),
+                                     color = "#F14628") +
+  ggtitle("Philly")+
+  theme(legend.position="none")
+
+NYscatter1 <- ggplot() +
+  geom_point(data = goodluck_ny, aes(x = miles, y = minutes),
+                                     color = "#2360a5") +
+  ggtitle("NYC")+
+  theme(legend.position="none")
+
+PHhisto2 <- ggplot() +
+  geom_histogram(data = goodluck_ph, aes(x = minutes),fill = "#F14628") +
+  ggtitle("Philly")+
+  theme(legend.position="none")
+
+NYhisto2 <- ggplot() +
+  geom_histogram(data = goodluck_ny, aes(x = minutes),fill = "#2360a5") +
+  ggtitle("NYC")+
+  theme(legend.position="none")
+
+```
+
+```{r , echo = FALSE, message= FALSE, fig.width=10, fig.height = 10}
+grid.arrange(DChisto, PHhisto2, NYhisto2,DCscatter, PHscatter1, NYscatter1, ncol = 3)
+
+```
+
+```{r echo = FALSE}
+acity_dat <- play_DC
+
+foo <- acity_dat %>%
+  arrange(minutes)
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_short_dc <- gg_dc +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#######
+#getting top 10% sample
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_long_dc <- gg_dc +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#some of the middle
+
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+#getting 10% of rows
+x <- (nrow(foo)*.1)*.5
+
+x <- (0.5*nrow(acity_dat))-x
+y <- (0.5*nrow(acity_dat))+x
+
+foo <- foo[x:y,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_med_dc <- gg_dc +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+```
+
+```{r echo = FALSE}
+
+acity_dat <- play_PH
+
+foo <- acity_dat %>%
+  arrange(minutes)
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_short_ph <- gg_ph +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#######
+#getting top 10% sample
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_long_ph <- gg_ph +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#some of the middle
+
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+#getting 10% of rows
+x <- (nrow(foo)*.1)*.5
+
+x <- (0.5*nrow(acity_dat))-x
+y <- (0.5*nrow(acity_dat))+x
+
+foo <- foo[x:y,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_med_ph <- gg_ph +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+```
+
+```{r echo = FALSE}
+
+acity_dat <- play_NY
+
+foo <- acity_dat %>%
+  arrange(minutes)
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_short_ny <- gg_ny +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#######
+#getting top 10% sample
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+
+#getting 10% of rows
+x <- nrow(foo)*.1
+
+foo <- foo[1:x,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_long_ny <- gg_ny +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+#some of the middle
+
+foo <- acity_dat %>%
+  arrange(desc(minutes))
+#getting 10% of rows
+x <- (nrow(foo)*.1)*.5
+
+x <- (0.5*nrow(acity_dat))-x
+y <- (0.5*nrow(acity_dat))+x
+
+foo <- foo[x:y,]%>%
+  sample_n(50)
+
+foopath <- foo %>%
+  select(latitude = endLat, longitude = endLon, trip_id) %>%
+  mutate(type = rep("destination"))
+foopath1 <- foo %>%
+  select(latitude = startLat, longitude = startLon, trip_id) %>%
+  mutate(type = rep("origin"))
+foopath <- rbind(foopath, foopath1)
+foopath$type <- as.factor(foopath$type) 
+foopath$trip_id <- as.factor(foopath$trip_id)
+
+along_med_ny <- gg_ny +
+  geom_point(data = foopath, aes(x = longitude, y = latitude),color = "#fdf5e6", size = 1.5)+
+  geom_path(data=foopath, aes(x=longitude, y=latitude, group=trip_id), color = "#fdf5e6", size=0.4)+
+  theme(legend.position="none")
+
+```
+
+### Sample of Short, Medium, and Long Trips {-}
+I wanted to see what some of the different length trips looked like in each city.
+Below is series of graphs showing short, medium, and long trips.
+This was done by taking sample of 50 trips from the bottom, middle, and top 10% of trips for each city.
+
+```{r echo = FALSE, fig.height = 10, fig.width = 12}
+grid.arrange(along_short_dc,along_med_dc, along_long_dc,
+             along_short_ph,along_med_ph, along_long_ph,
+             along_short_ny,along_med_ny, along_long_ny, ncol = 3)
+```
+
+##Neighborhoods {-}
+
+As a last question, I was wondering where within each city was the best origin. Below are the top ten origins.
+
+```{r echo = FALSE}
+##Philly
+
+foo <- play_PH %>%
+  group_by(startLat)%>%
+  filter(n()>2)%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
+head_neighborhoods_ph <- data.frame(
+        latitude = foo$startLat[1:10],
+        longitude =foo$startLon[1:10])
+
+along <- gg_ph +
+  geom_point(data = head_neighborhoods_ph, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)
+
+leaf_ph_neigh <- 
+  leaflet(philly) %>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addProviderTiles("Esri.WorldStreetMap") %>%
+  addCircleMarkers(lng = head_neighborhoods_ph$longitude, lat = head_neighborhoods_ph$latitude,
+                   radius = 2, color = "red")
+
+leaf_ph_neigh
+
+```
+
+<br>
+
+```{r echo = FALSE}
+###DC
+
+foo <- play_DC %>%
+  group_by(startLat)%>%
+  filter(n()>3 )%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
+head_neighborhoods_dc <- data.frame(
+  latitude = foo$startLat[1:10],
+  longitude =foo$startLon[1:10])
+
+along <- gg_dc +
+  geom_point(data = head_neighborhoods_dc, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)
+
+leaf_dc_neigh <- 
+  leaflet(dc) %>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addProviderTiles("Esri.WorldStreetMap") %>%
+  addCircleMarkers(lng = head_neighborhoods_dc$longitude, lat = head_neighborhoods_dc$latitude,
+                   radius = 2, color = "red")
+
+leaf_dc_neigh
+```
+
+<br>
+
+```{r echo = FALSE}
+###NY
+
+foo <- play_NY %>%
+  group_by(startLat)%>%
+  filter(n()>20 )%>%
+  summarise_all(mean) %>%
+  arrange(minutes)
+
+#the winner has 1 trip and is delaware rive to delaware river and takes ~20 minutes
+head_neighborhoods_ny <- data.frame(
+  latitude = foo$startLat[1:10],
+  longitude =foo$startLon[1:10])
+
+along <- gg_ny +
+  geom_point(data = head_neighborhoods_ny, aes(x = longitude, y = latitude), color = "#fdf5e6",  size = 1.5)
+
+leaf_ny_neigh <- 
+  leaflet(ny) %>%
+  addPolygons(fillColor = "#8a07ef", weight = 1, smoothFactor = 0.2,
+              opacity = 0.5 ,color = "white")%>%
+  addProviderTiles("Esri.WorldStreetMap") %>%
+  addCircleMarkers(lng = head_neighborhoods_ny$longitude, lat = head_neighborhoods_ny$latitude,
+                   radius = 2, color = "red")
+
+leaf_ny_neigh
+```
+
+# Endnotes:
+
+**************
+[^1]: One of only 3 positions in life that take shotgun seat in front of good friend. The other two being parents and grandparents.
+
+[^2]: This made the trip much more annoying.
+In Philadelphia they had purchased a new caspar mattress during a 4th of July sale.
+My friend had done this forgetting he'd offered me a ride...the mattress and I were quite squished in back seat.
+
+[^3]: The reason for this is purely personal.
+I had spent the best 10 years of my childhood around Philadelphia.
+The three of us in the car had attended University of Maryland, and so had used DC public transportation regularly over the last 4 years.
+Lastly, my friends were two weeks from moving to NYC to start their lives together.
+
+[^4]: [Wikipedia ShapefileArticle](https://en.wikipedia.org/wiki/Shapefile)
+
+[^5]: I hope it's clear enough from the code chunk, but I want to make it clear that the numextract function is not my own. I found it on: http://stla.github.io/stlapblog/posts/Numextract.html
+
+[^6]: Spoilder alert: not for long.
+
+[^7]: status codes starting 3xx or 4xx spell out trouble. Also, can use the httr function `status_code()` to get this piece of information easier.
+
+[^8]: This includes some absurd areas like within rivers and oceans.
+[^9]: I used the [HERE API](https://developer.here.com/)

### 2019-01-02 content/post/2019-01-02-transit.Rmd

@@
 tags: ["R Markdown", "plot", "regression"]
-output: 
-  html_document:
-    self_contained: no
+
 ---
@@
 library(rowr)
+library(png)
+library(knitr)
 
@@
 ```{r echo=FALSE, out.width = '100%'}
-knitr::include_graphics("x1.png")
-
-```
+
+img1_path <- "data/transit/x2.png"
+include_graphics(img1_path)
+```
+
 ```{r echo=FALSE, out.width = '35%'}
-knitr::include_graphics("Untitled.png")
+
+img2_path <- "data/transit/Untitled.png"
+include_graphics(img2_path)
 ```
@@
 ```{r echo=FALSE, out.width = '50%'}
-knitr::include_graphics("Septa_logo.png")
+
+img3_path <- "data/transit/Septa_logo.png"
+include_graphics(img3_path)
 
@@
 ```{r echo=FALSE, out.width = '25%'}
-knitr::include_graphics("mta_logo.png")
+
+img4_path <- "data/transit/mta_logo.png"
+include_graphics(img4_path)
 
@@
 ```{r echo=FALSE, out.width = '25%'}
-knitr::include_graphics("wmata_logo.png")
+
+img5_path <- "data/transit/wmata_logo.png"
+include_graphics(img5_path)
 

### 2019-01-03 content/post/article.Rmd

---
title: "Airbnb Predictions Project"
author: "Effi Feldblum"
date: '2018-12-10'
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]

---

# Introduction

This work  was originally done for a university project. The project had two main components: a kaggle-esque competition and a full business report. Due to this dual mandate, there are numerous methods attempted that would be omitted. The competition was judged based off of plain accuracy of a held-out testing dataset. This document is a sort of summary of the finalized pipeline. I hope it is helpful to any of you out there.

This is an analysis of the factors that potentially influence Airbnb booking numbers in an attempt to predict which listings will be highly active. Highly active is defined in the binary high_booking_rate variable with 1 indicating a listing is booked more than 90% of the available time. The proportion of highly booked listings is relatively small at ~25%. There are two main goals of the analysis. First, to build a model reasonably more effective at classifying listings as high booking rate than the baseline. Second, to understand which factors are most important in identifying highly booked listings in order to provide hosts with information.

## Setup {.tabset}

### Libraries
```{r Libraries, message=FALSE, warning=FALSE}
library(DataExplorer)
library(tidyverse)
library(data.table)
library(caret)
library(DMwR)
library(purrr)
library(gridExtra)
library(h2o)
library(kableExtra)
#library(glmnet) Can be really annoying with mlr's auc metric. Not sure why or how to fix besides unloading it before using mlr.
```

### Functions used
```{r Functions}
missing_plot <- function(df){
  xx <- colSums(is.na(DF_Test))
  xx1 <- as.data.frame(xx)
  rws <- row.names(xx1)
  xx1 <- xx1 %>%
    dplyr::mutate(percent = round((xx/nrow(DF_Test))*100),4)%>%
    mutate(variable = rws)%>%
    arrange(xx)

  ggplot(aes(x= variable, y =xx, label = percent), data = xx1)+
    geom_col()+
    coord_flip()

}

find_rate_cleaning <- function(amount){
  #filter out by the amount requessted
  xx = filter(DF_Test,cleaning_fee >amount)
  #basically a table like thing
  xx=count(xx,high_booking_rate)
  return(xx[2,2]/(xx[1,2]+xx[2,2]))
}

find_rate_cleaning_opposite <- function(amount){
  #filter out by the amount requessted
  xx = filter(DF_Test,cleaning_fee <amount)
  #basically a table like thing
  xx=count(xx,high_booking_rate)
  return(xx[2,2]/(xx[1,2]+xx[2,2]))
}

```

# EDA

Let's read in the data and convert all the character variables to factors.

```{r Data Load}
DF_Test <- read.csv("data/airbnb/Airbnb_Training.csv", stringsAsFactors = FALSE,na.strings=c("","NA"))
backup <- DF_Test
DF_Test$high_booking_rate <- as.factor(DF_Test$high_booking_rate)
levels(DF_Test$high_booking_rate) <- c("Low", "High")

fact_col <- colnames(DF_Test)[sapply(DF_Test,is.character)]
for(i in fact_col)
  set(DF_Test,j=i,value = factor(DF_Test[[i]]))

glimpse(DF_Test)
```
So, there a total of `r ncol(DF_Test)` columns and  `r nrow(DF_Test)` rows. The dependent variable is high booking rate.

One of the first things noticed about the high booking rate was the relative class imbalance, with around 75% of the observations in the full dataset having a low booking rate. Booking rate is defined as high (or 1, as a dummy variable) if the listing is booked more than 90% of the available time, and low (or 0) otherwise. As we are attempting to identify and predict Airbnb listings that will have a high booking rate, there was considerable effort to understand the reasons behind this class imbalance, and combat where possible.

```{r}
table(DF_Test$high_booking_rate)
``` 
Another quick item was that noticed was that while host_total_listing_count supposedly contains listings that a host used to have, it is identical to the host_listing_count, so it was removed.

```{r Identical vs }

# checking for identical variables
identical(DF_Test$host_total_listings_count,DF_Test$host_listings_count)

```

## First Look

Below is a condensed set of charts. As you can see, there are heavy skews and issues with deficient classes.

```{r EDA1, warning=FALSE}
#From the DataExplorer package
plot_histogram(DF_Test)
```
These histograms show some worrying skew and potential outliers in some of our numerical variables. This was dealt with below with some identification of faulty inputs and full transformations on variables.

```{r EDA2, warning=FALSE}
plot_bar(DF_Test)
```
Some variables that will need attention here are the market, property type, bed type. All have either some abnormalities or class deficiency.

<br>
<br>

The main approach was to transform the dataset into two main variations: one more heavily manipulated and variables were reduced (Modified), and the other was kept more to the original (Lean). Then, several transformations, factor level decisions, and models were tested.

## Missing Values
```{r missing values}
# quick custom function, can be found below
missing_plot(DF_Test)

colSums(is.na(DF_Test))

```

Okay so there's a fair number  of missing values, ~ 2.5% of all observations. These missing values were split among several variables.  Treatment  was experimented with and our final pipeline contained different methods.

### cleaning fees

The two variables with the largest proportion of missing variables were cleaning_fee and security_deposit. The main question is whether these NAs are legitimate, there on purpose, or due to a scraping or entry  error.

```{r cleaning_fees}

summary(DF_Test$cleaning_fee)
nrow(filter(DF_Test, cleaning_fee <25)) # there are 7000ish below 25
nrow(filter(DF_Test, cleaning_fee ==0)) # there are 767 at 0 (There are double the number at $10 cleaning fee)

```
The average cleaning fee is not absurd; however, the indicated skew is quite severe. Using a combination of domain knowledge and this data, it was suspected that these NAs were likely due to that listing lacking those charges. From experience, there are many listings without either a cleaning fee or a security deposit. When looking at cleaning fee, ~60% are beneath \$100, yet there are only 1.7% set at zero (there are 5% set exactly at $25). The story is much the same for security_deposit NAs. Because of this, it was decided to impute all NAs as 0. 
To take a quick peak at the skew.
```{r cleaning_fees1, message =  FALSE, warning=FALSE}
ggplot() +geom_histogram(aes(x=cleaning_fee), data = DF_Test)
#as seen above the max is 1000$ cleaning...e
#so there's def a skew here
nrow(filter(DF_Test, cleaning_fee >300)) # there are 159 listings above $300
nrow(filter(DF_Test, cleaning_fee >400)) # only 15 listings above $500

```

```{r cleaning_fee2}

c <- c(5,10,25,50,100,200,300,400)
sapply(c, find_rate_cleaning) # a custom function that finds the high booking rate of all observations above that number

sapply(c,find_rate_cleaning_opposite)# a custom function that finds the high booking rate of all observations above that number
```

What can be seen here is mirrored in many of the skewed numerical variables. There are several listings in either mansions or famous residences with prices and descriptions that indicate a high booking rate might not be necessary for their business. However, if you're looking to have a high booking rate, there appears to be an incentive to remain "sane" with your added fees. Perhaps there should  be some testing with a cleaning_fee per bedroom or per accomodates.

# Dataset Modification{.tabset}

## Description

The first tab is the setup for trying different levels for market with three different binnings attempted.
The next two tabs are used to create the bases for the two datasets referenced above [Lean, Modified].

## Market Setup

Markets
```{r fig.height=2.5, fig.width=2.5}
DF_market_18 <- DF_Test
DF_market_18$market <- as.character(DF_market_18$market)

DF_market_18$market <- ifelse(DF_market_18$market == "Nice", "Other (International)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Venice", "Other (International)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "London", "Other (International)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Agra", "Other (International)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Lagos, NG", "Other (International)",DF_market_18$market)

DF_market_18$market <- ifelse(DF_market_18$market == "San Antonio, US", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Philadelphia", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Dallas", "Other (Domestic)",DF_market_18$market)

DF_market_18$market <- ifelse(DF_market_18$market == "Dallas", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "San Antonio, US", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Indianapolis", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Providence", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Fresno", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Houston", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Fontana", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "South Florida Gulf Coast", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Fontana", "Other (Domestic)",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Pittsburg", "Other (Domestic)",DF_market_18$market)

DF_market_18$market <- ifelse(DF_market_18$market == "East Bay, CA", "San Francisco",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "South Bay, CA", "Los Angeles",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Coastal Orange County", "Los Angeles",DF_market_18$market)
DF_market_18$market <- ifelse(DF_market_18$market == "Carlsbad", "San Diego",DF_market_18$market)
DF_market_18$market <- as.factor(DF_market_18$market)

DF_market_4 <- DF_market_18
DF_market_4$market <- as.character(DF_market_4$market)

DF_market_4$market <- ifelse(DF_market_4$market == "New Orleans", "Mid City",DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Austin", "Mid City",DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "D.C.", "Mid City",DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "New Orleans", "Mid City",DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Chicago", "Mid City" ,DF_market_4$market)

DF_market_4$market <- ifelse(DF_market_4$market == "Nashville", "Mid City" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Portland", "Mid City" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Boston", "Mid City" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Denver", "Mid City" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Seattle", "Mid City" ,DF_market_4$market)

DF_market_4$market <- ifelse(DF_market_4$market == "Los Angeles", "Cali" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "San Francisco", "Cali" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "San Diego", "Cali" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Malibu", "Cali" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "Monterey Region", "Cali" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market == "North Carolina Mountains", "Other (Domestic)" ,DF_market_4$market)
DF_market_4$market <- ifelse(is.na(DF_market_4$market), "Other (Domestic)" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market== "Other (Domestic)","Other" ,DF_market_4$market)
DF_market_4$market <- ifelse(DF_market_4$market== "Other (International)","Other" ,DF_market_4$market)
DF_market_4$market <- as.factor(DF_market_4$market)

# 2 markets, NYC and everywhere else

DF_market_2 <- DF_Test
DF_market_2$market <- as.character(DF_market_2$market)

DF_market_2$market <- ifelse(DF_market_2$market == "New York", "New York","Out of Town") #send ben a screen shot

DF_market_2$market <- as.factor(DF_market_2$market)

plot_bar(DF_market_2$market)
plot_bar(DF_market_4$market)
plot_bar(DF_market_18$market)

```

Properties
```{r}

DF_prop_5 <- DF_Test

house <- c("Boat", "Chalet", "House", "Guesthouse", "Townhouse", "Vacation home", "Villa", "Tiny house", "Cottage","Bungalow" )
apartment <- c("Guest suite","apartment", "Apartment", "Loft", "Condominium","In-law","Serviced apartment","Dorm")
hotel <-  c("Resort", "Aparthotel","Bed & Breakfast", "Bed and breakfast","Boutique hotel","Hotel", "Hostel","Dorm")
nature <- c("Treehouse", "Cabin",  "Camper/RV",  "Earth House", "Earth house", "Tent", "Hut", "Yurt", "Tipi", "Nature Lodging", "Cave","Nature lodge")
other <- c("Other", "Castle", "Plane", "Barn", "Timeshare",NA)

DF_prop_5$property_type <- as.character(DF_prop_5$property_type)
DF_prop_5$property_type <- ifelse(DF_prop_5$property_type %in% house, "house",DF_prop_5$property_type)
DF_prop_5$property_type <- ifelse(DF_prop_5$property_type %in% apartment, "apartment",DF_prop_5$property_type)
DF_prop_5$property_type <- ifelse(DF_prop_5$property_type %in% hotel, "hotel",DF_prop_5$property_type)
DF_prop_5$property_type <- ifelse(DF_prop_5$property_type %in% other, "other",DF_prop_5$property_type)
DF_prop_5$property_type <- ifelse(DF_prop_5$property_type %in% nature, "nature",DF_prop_5$property_type)

DF_prop_5$property_type <- as.factor(DF_prop_5$property_type)
rm(list = c("house", "apartment", "hotel", "nature", "other"))
```

```{r echo=FALSE, fig.height=3, fig.width=3}
plot_bar(DF_market_2$market)
plot_bar(DF_market_4$market)
plot_bar(DF_market_18$market)
```

We can also take a peak at the property categories we created:

```{r echo=FALSE, fig.height=3, fig.width=3}
plot_bar(DF_prop_5$property_type)
```

Okay, based on that we're going to recombine the last three categories into other.

```{r}
DF_prop_2 <- DF_prop_5
DF_prop_2$property_type <- as.character(DF_prop_2$property_type) 
DF_prop_2$property_type <- ifelse(DF_prop_2$property_type=="hotel", "other", DF_prop_2$property_type)
DF_prop_2$property_type <- ifelse(DF_prop_2$property_type=="nature", "other", DF_prop_2$property_type)
DF_prop_2$property_type <- as.factor(DF_prop_2$property_type)
```

## Lean Dataset

```{r}

DF.lean <- DF_Test %>%
  dplyr::select(-host_total_listings_count, -amenities, -description, -host_verifications, -summary)
DF.lean$host_listings_count[DF.lean$host_listings_count==""] <- 0
DF.lean$bathrooms[is.na(DF.lean$bathrooms)] <- 0
DF.lean$beds[is.na(DF.lean$beds)] <- 0
DF.lean$bedrooms[is.na(DF.lean$bedrooms)] <- 0
DF.lean$host_has_profile_pic[is.na(DF.lean$host_has_profile_pic)] <- 'f'
DF.lean$host_identity_verified[is.na(DF.lean$host_identity_verified)] <- 'f'
DF.lean$host_is_superhost[is.na(DF.lean$host_is_superhost)] <- 'f'
DF.lean$host_listings_count[is.na(DF.lean$host_listings_count)] <- 0
DF.lean$security_deposit[is.na(DF.lean$security_deposit)] <- 0
DF.lean$cleaning_fee[is.na(DF.lean$cleaning_fee)] <- 0
DF.lean$host_response_rate[is.na(DF.lean$host_response_rate)] <- 0

DF.lean$host_response_time <- as.character(DF.lean$host_response_time)
DF.lean$host_response_time[is.na(DF.lean$host_response_time)] <- "unknown"
DF.lean$host_response_time <- as.factor(DF.lean$host_response_time)

#######WHere to add market type and property type

## DF 1 for different market types
DF.lean <- DF.lean%>% dplyr::select(-property_type, -market)
DF.lean <- cbind(DF.lean, DF_market_2$market, DF_prop_5$property_type)

DF.lean <- DF.lean%>%
  rename(market = `DF_market_2$market` )%>%
  rename(property_type = `DF_prop_5$property_type`)

DF.lean$market <- as.character(DF.lean$market)
DF.lean$market[is.na(DF.lean$market)] <- "unknown"
DF.lean$market <- as.factor(DF.lean$market)

DF.lean$property_type <- as.character(DF.lean$property_type)
DF.lean$property_type[is.na(DF.lean$property_type)] <- "unknown"
DF.lean$property_type <- as.factor(DF.lean$property_type)

DF.lean$cancellation_policy <- factor(DF.lean$cancellation_policy)
DF.lean$host_identity_verified<- factor(DF.lean$host_identity_verified)
DF.lean$host_is_superhost<- factor(DF.lean$host_is_superhost)
DF.lean$host_response_time<- factor(DF.lean$host_response_time)
DF.lean$instant_bookable<- factor(DF.lean$instant_bookable)

#only numeric variables
DF.leana <- DF.lean[ , purrr::map_lgl(DF.lean, is.numeric)]

#remove the listing id into seperate objects and delete
listing_ids <- DF.lean$Listing_ID
target_class <- DF.lean$high_booking_rate
DF.lean <- DF.lean %>%
  dplyr::select(-high_booking_rate, -Listing_ID)

#making the weird max nights above 1825 to 1825
DF.lean$maximum_nights <- ifelse(DF.lean$maximum_nights>1825, 1825, DF.lean$maximum_nights)
#making beds real or non real bed
DF.lean$bed_type <- as.character(DF.lean$bed_type)
DF.lean$bed_type <- ifelse(DF.lean$bed_type=="Real Bed", 1,0)
DF.lean$bed_type <- as.factor(DF.lean$bed_type)

# Add new variable for beds per bedroom
DF.lean$beds_per_bedroom <- ifelse(DF.lean$bedrooms != 0, DF.lean$beds/DF.lean$bedrooms, DF.lean$beds)

# Add new variables for availabilities
DF.lean$availability_3060 <- DF.lean$availability_60 - DF.lean$availability_30
DF.lean$availability_6090 <- DF.lean$availability_90 - DF.lean$availability_60

DF.lean <- DF.lean %>% dplyr::select(-availability_365, -availability_60, -availability_90,-availability_30,-latitude, -longitude)

DF.lean <- cbind(target_class, DF.lean)
DF.lean <- DF.lean %>% rename(high_booking_rate = target_class)

DF.lean$cancellation_policy <- as.character(DF.lean$cancellation_policy)
DF.lean <- DF.lean %>%
  filter(cancellation_policy != "no_refunds") #because of issues with test data
DF.lean$cancellation_policy <- as.factor(DF.lean$cancellation_policy)

levels(DF.lean$high_booking_rate) <- c("0", "1")

```

## Modified Dataset

```{r}

DF.1 <- backup %>%
  dplyr::select(-host_total_listings_count,-market)

DF.1$security_deposit[is.na(DF.1$security_deposit)] <- 0
DF.1$cleaning_fee[is.na(DF.1$cleaning_fee)] <- 0

DF_categ <- DF.1[ , purrr::map_lgl(DF.1, is.character)]
rwww <- colnames(DF_categ)
DF_categ <- DF_categ %>% dplyr::select(-summary, -description)
DF_categ[sapply(DF_categ, is.character)] <- lapply(DF_categ[sapply(DF_categ, is.character)], 
                                                   as.factor)

DF.2 <- dplyr::select(DF.1 , -rwww)
DF.1 <- cbind(DF.2,DF_categ)

# Add new variable for beds per bedroom
DF.1$beds_per_bedroom <- ifelse(DF.1$bedrooms != 0, DF.1$beds/DF.1$bedrooms, DF.1$beds)

#Might want to comment this out # Add new variables for availabilities    
 DF.1$availability_3060 <- DF.1$availability_60 - DF.1$availability_30
 DF.1$availability_6090 <- DF.1$availability_90 - DF.1$availability_60
 DF.1 <- DF.1 %>%
   dplyr::select(-availability_30, -availability_365, -availability_60,-availability_90)

DF.1 <- DF.1 %>%
  dplyr::select(-latitude, -longitude, - amenities, -Listing_ID,-host_verifications,-property_type)

#######WHere to add market type and property type

## DF 1 for different market types
DF.1.m2 <- cbind(DF.1, DF_market_2$market, DF_prop_5$property_type)

DF.1.m2 <- DF.1.m2%>%
  rename(market = `DF_market_2$market` )%>%
  rename(property_type = `DF_prop_5$property_type`)

DF.1.m2$high_booking_rate <- factor(DF.1.m2$high_booking_rate)

#adding levels of NA
DF.1.m2$host_response_time <-addNA(DF.1.m2$host_response_time)

DF.1.m2$host_response_rate[is.na(DF.1.m2$host_response_rate)] <- 0

DF.1.m2$cancellation_policy <- as.character(DF.1.m2$cancellation_policy)
DF.1.m2$cancellation_policy <- ifelse(DF.1.m2$cancellation_policy == "no_refunds","super_strict_60",DF.1.m2$cancellation_policy)
DF.1.m2$cancellation_policy <- as.factor(DF.1.m2$cancellation_policy)

#basic KNN Imputation

knnOutput <- DMwR::knnImputation(DF.1.m2[, !names(DF.1.m2) %in% "high_booking_rate"])  # perform knn imputation.

DF.1.m2 <- cbind(DF.1.m2$high_booking_rate, knnOutput)

DF.1.m2 <- rename(DF.1.m2, high_booking_rate=`DF.1.m2$high_booking_rate`)

```

# Finding Best Datasets
So, the first thing done was to discover which transformations worked the best. Both datasets had the following transformations created:  
  * Center  
  * Scale  
  * Center & Scale  
  * BoxCox  
  * BoxCox & Center & Scale  
  * Yeo Johnson

Then, each dataset was run through an ordinary logistical regression and two random forests. Both random forests ran with 5 folds. The first balanced classes and used a modulo fold assignment while the second does a pure stratified sampling.

The code and results are below.

## Operation

Complete the transformation
```{r}
foo <- DF.lean

target_class_train <- foo$high_booking_rate
foo <- foo %>% select(-high_booking_rate)

preProcValues <- preProcess(foo, method= c("YeoJohnson"))
trainTransformed <- predict(preProcValues, foo)

#recomb
foo <- cbind(target_class_train,trainTransformed)
foo <- foo %>% rename(high_booking_rate = target_class_train)

```

The logistical regression.
```{r eval = FALSE}
datafr <- DF.lean

set.seed(9444)
trainIndex <- createDataPartition(datafr$high_booking_rate, p = .8,
                                  list = FALSE, 
                                  times = 1)
training <- datafr[trainIndex,]
validation <- datafr[-trainIndex,]

full_log <- glm(high_booking_rate~., data=training, family="binomial")
full_log_preds <- predict(full_log, newdata=validation, type="response")
full_log_class <- ifelse(full_log_preds>0.5,1,0)

table.lean <- table(validation$high_booking_rate, full_log_class)
confusionMatrix(table.lean)
```

The two random forests.
```{r message=FALSE, eval = FALSE, eval =FALSE}

localH2O <- h2o.init(nthreads = -1)
y <-  1 #that is the number it is, likely 1 for me
x <- c(2:30)

datafr.h20 <- as.h2o(datafr)

nfolds=5
splits <- h2o.splitFrame(data = datafr.h20, 
                         ratios = c(0.7, 0.15),  #partition data into 70%, 15%, 15% chunks
                         seed = 1)  #setting a seed will guarantee reproducibility

train <- splits[[1]]
valid <- splits[[2]]
test <- splits[[3]]
my_rf_mod <- h2o.randomForest(x = x,
                              y = y,
                              training_frame = train,
                              nfolds = nfolds,
                              balance_classes = TRUE,
                              fold_assignment = "Modulo",
                              keep_cross_validation_predictions = TRUE,
                              seed = 1)

my_rf_strat <- h2o.randomForest(x = x,
                                y = y,
                                training_frame = train,
                                nfolds = nfolds,
                                fold_assignment = "Stratified",
                                keep_cross_validation_predictions = TRUE,
                                seed = 1)

h2o.performance(my_rf_mod,test)
h2o.performance(my_rf_strat,test)

h2o.shutdown(prompt = TRUE)
```

The results:

```{r results1, echo = FALSE}
results1 <- data.frame(
  model = c("Modified","Modified Center" , "Modified Scale",
    "Modified Center Scale", "Modified BoxCox", "Modified Center Scale Box",
    "Lean", "Lean Center", "Lean Scale",  
    "Lean BoxCox", "Lean Scale center Boxcox","Lean YeoJohnson","Modified YeoJohnson"),

  logit_results = c(0.7841,0.7841,0.7841,0.7841,0.7867, 0.7867, 0.7885, 0.7885,
                    0.7885, 0.7888, 0.7888,0.7888, 0.7872),
  rf_modulo = c(0.7896084, 0.7841833, 0.8010347, 0.798494, 0.7948263, 0.7960089, 0.795861,
    0.7994087,  0.7878788, 0.7920177, 0.798374, 0.7900961, 0.8029564),
  rf_stratified =c(0.7924612, 0.7738359, 0.7841833, 0.7920177, 0.7945307, 0.7825573, 0.795122,
    0.7824095, 0.7955654, 0.8005913, 0.8002956, 0.8097561, 0.7899483))

results1 %>%
  kable() %>%
  kable_styling()
```

next step is determining which market and property levels to use.

## market level determination
Now we'll determine which of the market determinations to use. It was quickly seen that the property type with 5 categories was yielding better results, in terms of accuracy. The 12 best datasets above were varied with the three different market level types and tested the the same way as above.

```{r  echo = FALSE}

results_market <- data.frame(
  model = c("Lean", "Lean BoxCox", "Lean YeoJohnson", "Modified", 
            "Modified Scale",  "Modified YeoJohnson","Lean", "Lean BoxCox", "Lean YeoJohnson", "Modified", 
            "Modified Scale",  "Modified YeoJohnson", "Lean", "Lean BoxCox",
            "Lean YeoJohnson", "Modified", "Modified Scale",  "Modified YeoJohnson"),
  market_type = c(2,2,2,2,2,2,4,4,4,4,4,4,18,18,18,18,18,18),
  logit_results = c(0.7896,  0.791, 0.7929, 0.7858, 0.7858,  0.7893
                    ,0.7892, 0.7924, 0.7919, 0.7843, 0.7843, 0.7875,
                    0.7896, 0.791,  0.7929,0.7858, 0.7858, 0.7893   ),
  rf_modulo = c(0.8, 0.7974871, 0.8031042, 0.8094605,  0.8010347, 0.8029564
                ,0.7963045, 0.8001478, 0.804878, 0.7821138, 0.8022173, 0.7980783,
                0.8,  0.7974871, 0.8031042, 0.8094605,  0.8022173, 0.7980783  ),
  rf_stratified =c(0.8088692, 0.807391, 0.8031042, 0.8094605, 0.8022173, 0.7980783
            ,0.7918699, 0.7831486, 0.7920177, 0.7920177, 0.7927568,  0.8004435,
                   0.8088692, 0.807391, 0.7961567,  0.7982262, 0.7881744, 0.7974871  )
)
results_market %>%
  kable() %>%
  kable_styling()
```

# Modeling on the best 4 datasets{.tabset}
## Summary of Modeling
This was, admittely, a bit of throwing different models at the training data.Being for a class, there was an attempt to demonstrate and ability to code all of these. Furthermore, it being organized like a kaggle competition with points on the line, why not give it a try.

Here are the results, code can be seen in the other tab.

```{r echo=FALSE}

model_results <- data.frame(
  datasets = c("Modified 18 Markets", "Lean 18 Markets", "Lean 2 Markets YeoJohnson", "Modified 4 markets YeoJohnson"),
  ridge_cv = c(0.7818994, 0.7822627, 0.7912786,  0.7912786),
  lasso_cv = c( 0.7890783, 0.7892643, 0.7909675,0.7925233),
  LDA =c( 0.7712, 0.7764, 0.7888, 0.7888),
  `SVM Linear`=c(0.7846,0.7912, 0.7912, 0.7906 ),
  `SVM Radial` = c(0.81, 0.8105,0.8157, 0.8161),
  `Gradient Boosting Machine` = c(0.8097561,0.8116778, 0.8031042,0.8121212),
  Xgboost =c( 0.83,0.8248, 0.8226, 0.8273 ))

model_results %>%
  kable() %>%
  kable_styling()

```

## Code for Modeling

All data was partitioned into 80% training data and 20% validation data. This was deemed a reasonable split given both the amount of data and the relative size of the testing data (the validation contains ~8000 with testing at ~7000). Except for the GBMs which used an explicit validation dataset in tuning.

FOr both Ridge and Lasso, the model is completed on the entire dataset in a matrix form. This is done here.

```{r eval = FALSE}

dfram <- DF.1.m4.YeoJohnson #swapped through datasets here

allData <- model.matrix (~ .-1, dfram[,c(1:30)])
allData_X <- model.matrix( ~ .-1, dfram[,c(2:30)])

```

Ridge Regression
```{r  eval = FALSE}

ridge = cv.glmnet(as.matrix(allData_X), dfram$high_booking_rate, family="binomial", alpha=0)
best.lambda=ridge$lambda.min
ridge_probs = predict(ridge, s=best.lambda, newx=allData_X, type="response")
ridge_class = ifelse(ridge_probs>0.45,1,0) #Used a 45% cutoff

## Calculate accuracy:
sum(ifelse(ridge_class==dfram$high_booking_rate,1,0))/nrow(dfram)
```

Lasso 
```{r eval = FALSE}
lasso = cv.glmnet(as.matrix(allData_X), dfram$high_booking_rate, family="binomial", alpha=1)
best.lambda2=lasso$lambda.min
lasso_probs = predict(lasso, s=best.lambda2, newx=allData_X, type="response")
lasso_class = ifelse(lasso_probs>0.5,1,0)

## Calculate accuracy:
sum(ifelse(lasso_class==dfram$high_booking_rate,1,0))/nrow(dfram)

```

Linear Discriminant Analysis
```{r eval = FALSE}

lda.fit <- lda(high_booking_rate~., data = training)

lda.pred <- predict(lda.fit, validation)
lda.class <- lda.pred$class
lda.pred <- lda.pred$posterior[,2]

log1 <- table(lda.class, validation$high_booking_rate)
confusionMatrix(log1)
```

Support Vector Machines: Linear and Radial Kernels
```{r eval = FALSE}

set.seed(9444)
trainIndex <- createDataPartition(datafr$high_booking_rate, p = .8,
                                  list = FALSE, 
                                  times = 1)
training <- datafr[trainIndex,]
validation <- datafr[-trainIndex,]

svm_Linear <- svm(high_booking_rate~.,data = training, method="C-classification", kernel="linear")

svm.pred <- predict(svm_Linear, validation, type="prediction")
mtab<-table(svm.pred,validation$high_booking_rate)
confusionMatrix(mtab)

svm_Rad <- svm(high_booking_rate~.,data = training,method="C-classification", kernel="radial",  probability = TRUE)

svm.pred <- predict(svm_Rad, validation, type="prediction")
mtab<-table(svm.pred,validation$high_booking_rate)
confusionMatrix(mtab)

```

GBM tuning
The parameter lists were adjusted for each dataset.
This strategy and code was found online.

```{r eval = FALSE}
#datafr[,1] <- NULL
localH2O <- h2o.init(nthreads = -1)

y <-  1 #that is the number it is, likely 1 for me
x <- c(2:30)

datafr.h20 <- as.h2o(datafr)

nfolds=5
splits <- h2o.splitFrame(data = datafr.h20, 
                         ratios = c(0.7, 0.15)  #partition data into 70%, 15%, 15% chunks
                        )  
train <- splits[[1]]
valid <- splits[[2]]
test <- splits[[3]]

#Cartesian grid of gbm
# GBM hyperparamters
h2o.table(train[y])

gbm_params1 <- list(learn_rate = c(0.01, 0.1),
                    max_depth = c(3, 10, 1),
                    sample_rate = c(0.8, 1.0),
                    col_sample_rate = c(0.2, 0.5, 1.0))

# Train and validate a grid of GBMs
gbm_grid1 <- h2o.grid("gbm", x = x, y = y,
                      grid_id = "gbm_grid1",
                      training_frame = train,
                      validation_frame = valid,
                      ntrees = 100,
                      seed = 1,
                      hyper_params = gbm_params1)

# Get the grid results, sorted by AUC
gbm_gridperf1 <- h2o.getGrid(grid_id = "gbm_grid1", 
                             sort_by = "auc", 
                             decreasing = TRUE)
print(gbm_gridperf1)
# Random Grid Search

# GBM hyperparamters
gbm_params2 <- list(learn_rate = seq(0.1, 1, 0.05),
                    max_depth = seq(3, 25, 1),
                    sample_rate = seq(0.5, 1.0, 0.1),
                    col_sample_rate = seq(0.1, 1.0, 0.1))
search_criteria2 <- list(strategy = "RandomDiscrete", 
                         max_models = 60)

# Train and validate a grid of GBMs
gbm_grid2 <- h2o.grid("gbm", x = x, y = y,
                      grid_id = "gbm_grid2",
                      training_frame = train,
                      validation_frame = valid,
                      ntrees = 100,
                      seed = 1,
                      hyper_params = gbm_params2,
                      search_criteria = search_criteria2)

gbm_gridperf2 <- h2o.getGrid(grid_id = "gbm_grid2", 
                             sort_by = "auc", 
                             decreasing = TRUE)
print(gbm_gridperf2)

###

gbm_params <- list(learn_rate = seq(0.05, 1, 0.01),  #updated
                   max_depth = seq(3, 13, 1),
                   sample_rate = seq(0.6, 1.0, 0.1),  #updated
                   col_sample_rate = seq(0.1, 0.8, 0.1))

search_criteria <- list(strategy = "RandomDiscrete", 
                        max_runtime_secs = 480)  #updated

gbm_grid <- h2o.grid("gbm", x = x, y = y,
                     grid_id = "gbm_grid4",
                     training_frame = train,
                     validation_frame = test,
                     ntrees = 250,
                     seed = 1,
                     hyper_params = gbm_params,
                     distribution ="bernoulli",
                     search_criteria = search_criteria)

gbm_gridperf <- h2o.getGrid(grid_id = "gbm_grid4", 
                            sort_by = "auc", 
                            decreasing = TRUE)
print(gbm_gridperf)

best_gbm_model_id <- gbm_gridperf1@model_ids[[1]]
best_gbm <- h2o.getModel(best_gbm_model_id)

#Test set prediction
best_gbm_perf <- h2o.performance(model = best_gbm, 
                                 newdata = test)
h2o.confusionMatrix(best_gbm_perf)

h2o.shutdown(prompt = TRUE)
```

Xgboost
```{r eval = FALSE}

#create tasks
setDF(train)
setDF(test)
traintask <- makeClassifTask(data = train,target = "high_booking_rate")
testtask <- makeClassifTask(data = test,target = "high_booking_rate")

#do one hot encoding
traintask <- createDummyFeatures(obj = traintask)
testtask <- createDummyFeatures(obj = testtask)

lrn <- makeLearner("classif.xgboost",predict.type = "prob")
lrn$par.vals <- list(
  objective="binary:logistic",
  eval_metric="error",
  nrounds=200L,
  eta=0.1
)

#set parameter space
params <- makeParamSet(
  makeDiscreteParam("booster",values = c("gbtree","gblinear")),
  makeIntegerParam("max_depth",lower = 3L,upper = 25L),
  makeNumericParam("min_child_weight",lower = 1L,upper = 10L),
  makeNumericParam("subsample",lower = 0.5,upper = 1),
  makeNumericParam("colsample_bytree",lower = 0.5,upper = 1)
)

#set resampling strategy
rdesc <- makeResampleDesc("CV",stratify = T,iters=5L)

#search strategy
ctrl <- makeTuneControlRandom(maxit = 5L, tune.threshold = TRUE)

#set parallel backend
library(parallel)
library(parallelMap)
parallelStartSocket(cpus = 5)

tic()
#parameter tuning
mytune1 <- tuneParams(learner = lrn
                     ,task = traintask
                     ,resampling = rdesc
                     ,measures = auc
                     ,par.set = params
                     ,control = ctrl
                     ,show.info = T)

toc()
#set hyperparameters
lrn_tune1 <- setHyperPars(lrn,par.vals = mytune1$x)

#train model
xgmodel1 <- mlr::train(learner = lrn_tune1,task = traintask)

#predict model
xgpred1 <- predict(xgmodel1,testtask)
confusionMatrix(xgpred1$data$response,xgpred1$data$truth)
```

# Conclusions
Overall, Xgboost was by far the best model, regardless of dataset. Submitting the best resulted in a validation accuracy of ~83% and got 83.21% on the held-out testing dataset. This was only 0.0035 percent off of the top submission.

# Trying SMOTE
To combat this, we tried several re-sampling techniques. The first was on the actual data where we performed SMOTE  (Synthetic Minority Over-sampling Technique) on the datasets. SMOTE takes the minority class and creates brand new observations through a KNN procedure. We used the default of 5 nearest neighbors. It is important to notice that SMOTE was only performed on the training set of the data and not the validation set. This was to ensure that validation results would simulate the results on true testing data.

SMOTE was tested on our best dataset with the best model (Modified 4 markets Yeo Johnson, Xgboost). It did not lead to any improvement. However, it did result in a slight improved parity in errors between the classes. Overall, it was not helpful in this scenario.

```{r eval = FALSE}
datafr <- DF.1.m2

names(datafr)
datafr[,1] <- NULL

set.seed(9444)
trainIndex <- createDataPartition(datafr$high_booking_rate, p = .8,
                                  list = FALSE, 
                                  times = 1)
train <- datafr[trainIndex,]
test <- datafr[-trainIndex,]

#create tasks
setDF(train)
setDF(test)
traintask <- makeClassifTask(data = train,target = "high_booking_rate")
testtask <- makeClassifTask(data = test,target = "high_booking_rate")

##
##What's DIFFERENT
task.smote = smote(traintask, rate = 3, nn = 5)

testtask <- createDummyFeatures(obj = testtask)
train.smote <- createDummyFeatures(obj = task.smote)
####

lrn <- makeLearner("classif.xgboost",predict.type = "prob")
lrn$par.vals <- list(
  objective="binary:logistic",
  eval_metric="error",
  nrounds=200L,
  eta=0.1
)

#set parameter space
params <- makeParamSet(
  makeDiscreteParam("booster",values = c("gbtree","gblinear")),
  makeIntegerParam("max_depth",lower = 3L,upper = 25L),
  makeNumericParam("min_child_weight",lower = 1L,upper = 10L),
  makeNumericParam("subsample",lower = 0.5,upper = 1),
  makeNumericParam("colsample_bytree",lower = 0.5,upper = 1)
)

#set resampling strategy
rdesc <- makeResampleDesc("CV",stratify = T,iters=5L)

#search strategy
ctrl <- makeTuneControlRandom(maxit = 5L, tune.threshold = TRUE)

#set parallel backend
#library(parallel)
#library(parallelMap)
parallelStartSocket(cpus = 5)

tic()
#parameter tuning
mytune1 <- tuneParams(learner = lrn
                     ,task = train.smote
                     ,resampling = rdesc
                     ,measures = auc
                     ,par.set = params
                     ,control = ctrl
                     ,show.info = T)

toc()
#set hyperparameters
lrn_tune1 <- setHyperPars(lrn,par.vals = mytune1$x)

#train model
xgmodel1 <- mlr::train(learner = lrn_tune1,task = train.smote)

#predict model
xgpred1 <- predict(xgmodel1,testtask)
confusionMatrix(xgpred1$data$response,xgpred1$data$truth)
```

### Addendum
NLP: Within NLP I also attempted several methods on both the amenities and description variables. On the description variable I used the text2vec, tm, and tidytext packages to create vocabularies, prune unused words from the vocab, and remove stopwords. Then, I tried two different sentiment analysis dictionaries, afinn and and nrc. Afinn gives a relative score of positive/negative, while nrc only classifies the word but into several emotions rather than only positive/negative. Neither set of these added variables added significant predictive poIr. The next thing tried was LDA (Latent Dirichlet Allocation), which is a type of topic modeling that uses the text to extract potential clusters prespecified. This method seemed to work best with three topics (they also seemed to be meaningful clusters); hoIver, even at its most effective, it did not add predictive poIr.

The last thing attempted was creating a non-pruned vocab of the amenities and splitting them into dummy variables I felt might be significant. It is possible that this might lead to significant improvement, but would require more analysis into individual variables. When all added variables that I believed would be “significant” (e.g. wifi, breakfast included, park) there was not much improvement, but some of the variables appeared to be significant.

A lot of this was attempted early on in the project and I'd like to come back and try it with some of the transformed datasets found above.

### 2019-01-03 public/post/2019-01-02-transit.Rmd

@@
 ---
-title: Transit
+title: A Question on the Efficiency of Transit in Absurd Situations
 author: Effi Feldblum
-date: '2019-01-02'
+date: '2018-08-22'
 categories: ["R"]
 tags: ["R Markdown", "plot", "regression"]
+
 ---
@@
 
-```{r echo=FALSE, out.width = '100%', eval = FALSE}
-knitr::include_graphics("google_maps_whats_here.png")
-
-```
-```{r echo=FALSE, out.width = '35%', eval = FALSE}
-knitr::include_graphics("Untitled.png")
+```{r echo=FALSE, out.width = '100%'}
+
+img1_path <- "data/transit/x2.png"
+include_graphics(img1_path)
+```
+
+```{r echo=FALSE, out.width = '35%'}
+
+img2_path <- "data/transit/Untitled.png"
+include_graphics(img2_path)
 ```
@@
 
-```{r echo=FALSE, out.width = '50%',eval = FALSE}
-knitr::include_graphics("Septa_logo.png")
+```{r echo=FALSE, out.width = '50%'}
+
+img3_path <- "data/transit/Septa_logo.png"
+include_graphics(img3_path)
 
@@
 
-```{r echo=FALSE, out.width = '25%',eval = FALSE}
-knitr::include_graphics("mta_logo.png")
+```{r echo=FALSE, out.width = '25%'}
+
+img4_path <- "data/transit/mta_logo.png"
+include_graphics(img4_path)
 
@@
 
-```{r echo=FALSE, out.width = '25%', eval = FALSE}
-knitr::include_graphics("wmata_logo.png")
+```{r echo=FALSE, out.width = '25%'}
+
+img5_path <- "data/transit/wmata_logo.png"
+include_graphics(img5_path)
 

### 2019-01-03 content/post/2019-01-02-transit.Rmd

@@
 ---
-title: Transit
+title: A Question on the Efficiency of Transit in Absurd Situations
 author: Effi Feldblum
-date: '2019-01-02'
+slug: transit
+description: A short Article
+summary: AAAA
+date: '2018-08-22'
 categories: ["R"]

### 2019-01-03 content/post/article.Rmd

@@
 ---
-title: "Airbnb Predictions Project"
-author: "Effi Feldblum"
+title: Airbnb Predictions Project
+author: Effi Feldblum
+slug: airbnb
+description: A short Article
+summary: AAAA
 date: '2018-12-10'

### 2019-01-03 content/post/article.Rmd

--- 2019-01-03 content/post/article.Rmd
@@
 ---
-title: "Airbnb Predictions Project"
-author: "Effi Feldblum"
+title: Airbnb Predictions Project
+author: Effi Feldblum
+slug: airbnb
+description: A short Article
 date: '2018-12-10'

### 2019-01-03 content/post/2019-01-02-transit.Rmd

--- 2019-01-02 content/post/2019-01-02-transit.Rmd
@@
 ---
-title: Transit
+title: A Question on the Efficiency of Transit in Absurd Situations
 author: Effi Feldblum
-date: '2019-01-02'
+slug: transit
+description: A short Article
+summary: AAAA
+date: '2018-08-22'
 categories: ["R"]
-tags: ["R Markdown", "plot", "regression"]
+tags: ["R", "AWS", "Parallel Computing", "GIS", "API use"]
 

### 2019-01-03 content/post/article.Rmd

--- 2019-01-03 content/post/article.Rmd
@@
 ---
-title: "Airbnb Predictions Project"
-author: "Effi Feldblum"
+title: Airbnb Predictions Project
+author: Effi Feldblum
+slug: airbnb
+description: A short Article
 date: '2018-12-10'
 categories: ["R"]
-tags: ["R Markdown", "plot", "regression"]
+tags: ["R", "Machine Learning", "Data Munging"]
 
